---
title: "bootstrap resampling of phacelia dataset: 5-5-2017"
author: "Jessie Berta-Thompson"
date: "May 5, 2017"
output: html_document
---

Yesterday's script is running Fst calculations on bootstrapped datasets separated by sampling location (15 categories): subsample_and_run_popgen_V1_5_4_17.Rmd

Copy much of that work here, and perform the same calculations for data sorted by species (3 categories): resample data, with replacement, with name number of samples as actually taken for each species, then run strataG Fst calculations on each.


*************************************************************************************
## Set up
Clear history and make sure you're in the right working directory for shorter file writes. 
Gather tools (reading and converting popgen data, running Fst calculations)
```{r}
rm(list=ls())
print("Where are we? Where will new files without specified full path be created?")
getwd()
library(ape)
library(phangorn)
library(ade4)
library(adegenet)
library(apex)
library(strataG)
```


*************************************************************************************
Step 1. Get the Phacelia microsatellite data

Read in the data (adegenet tools take a genepop.gen file, make a genind object)
```{r}
Real_data.genind <-read.genepop("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Subsampling/Copy_of_real_data_PhfoAllGenePop_13Dec2016.gen")

#Note one error message: "Duplicate individual names detected. Coercing them to be unique." This is because .gen file has no individual names, just a list of genotypes, but genind requires individual names. They are given numbers, in order of original file.
```

Take a quick look at the genind object
```{r}
print("Real data genind (by sampling population)")
#print on whole object gives an overview
print(Real_data.genind)

#number of samples per location
print(table(Real_data.genind$pop))

#the 12 loci and their alleles?
print(Real_data.genind$all.names)

#for each individual (rows), the number of each allele of each locus
print(Real_data.genind$tab[1:15,1:5])
```

This is one way of organizing the data - by sampling location, but today we will also be working at the species level, so  we need to map these site codes to the three putative species across the site map - copy of a file that maps sites to species lives here:
Q:\Research\All_Projects_by_Species\Phacelia SPECIES\Phacelia_formosula\Phacelia formosula_abiotic\Modelling\MetaPopGen\Copy_of_Species_data.xlsx

Gathering required information in this file here:
Q:\Research\All_Projects_by_Species\Phacelia SPECIES\Phacelia_formosula\Phacelia formosula_abiotic\Modelling\Subsampling\Summary_of_real_data.xlsx

Convert list of populations describing individuals to a list of species
```{r}
#grab that list of populations and convert to a vector for easy manipulation
pop_vector <- as.character(Real_data.genind@pop)

#how long?
print("population vector")
print(length(pop_vector))
data.class(pop_vector)
print(pop_vector)

#Make a new empty vector to hold the species version of this
species_vector <- character(length = length(pop_vector))

#make three character vectors with the mapping between species and populations
scullyi <- as.character(c("HN", "HS", "F"))
formosula <- as.character(c("CG", "D", "P","RN", "WS", "WA", "B", "V", "R", "AN", "AS"))
gina_glenneae <- as.character(c("T"))

#Loop through and perform mapping
for (i in 1:length(pop_vector)) {
  p <- pop_vector[i]
  if (p %in% scullyi) {
    species_vector[i] <- "s"}
  else if (p %in% formosula) {
    species_vector[i] <- "f"}
  else if (p %in% gina_glenneae) {
    species_vector[i] <- "g"}
  else {
    print("uhoh, population not found in library")
  }
}
#how long?
print("species vector")
print(length(species_vector))
data.class(species_vector)
print(species_vector)

#Make copies of the genind, and swap out the @pop factor for the species level designations. 
real_data_by_loc.genind <- Real_data.genind #more accurate name
real_data_by_species.genind <- Real_data.genind

real_data_by_species.genind@pop <- factor(species_vector)

#Species-based genind
print("Species-based genind")
print(real_data_by_species.genind)
print(table(real_data_by_species.genind$pop))

```


Build the bootstrapped dataset by resampling from the data organized by species
```{r}
#how many datasets to build?
bootreps <- 1000

#list with species designation for each individual
poplist <- real_data_by_species.genind$pop

#numeric vector with individual counts for each species
popcounts <- as.numeric(table(poplist))

#character vector - just the species names/designations
popnames <- names(table(poplist))

#how many species categories are there?
npop <- length(popnames)

#Create the bootstrap datasets
onethousanddatasets_by_species <- lapply(1:bootreps, function(bootrep) {
  
#convert genind to dataframe, where the rows are individuals, the columns are loci, and the entries are alleles.
ind_locus_table.df <- genind2df(real_data_by_species.genind, sep="-", pop = real_data_by_species.genind$pop, usepop = FALSE)

#the population assignment data lives in poplist object defined above, not in this dataframe.

#perform randome resampling, for each category (species)
listresults <- lapply(1:npop, function(p){
  name <- popnames[p] #which category are we on?
  n <- popcounts[p] #how many to sample for this category?
  #sample: only from the right category, same number as originally sampled for that category, 
  #sample at random with replacement (same individual can be sampled multiple times)
  resampled_rows <- sample(which(poplist== name), n, replace = TRUE) #specifically, which rows to take? 
  #take randomized rows from the data table and from population list in tandem (need to match; could also recreate pops from sampling scheme itself, which is by species)
  subsetdata <- ind_locus_table.df[resampled_rows,]
  subsetpops <- poplist[resampled_rows]
  
  #for each category, returns a list with 2 things. 
  #first is a df of data, second is a character vector with pop assignments
  return(list(subsetdata, subsetpops))
  })

#extract output of lapply list of lists results [[species 1 alleles, species 1 pop assignments], ...]
resampleddatalist <- lapply(listresults, function(sublist) {
  sublist[[1]]}) #creates a list of dataframes

resampledpoplist <- lapply(listresults, function(sublist) {
  sublist[[2]]})#creates a list of character vectors

#put different categories sampled back into a single dataframe
resampled.df <- do.call("rbind", resampleddatalist)
# put different pop assignments into a single list in order
resampledpops <- unlist(resampledpoplist)

#put back into a genind.
reconvert.genind <- df2genind(resampled.df, sep="-", pop = resampledpops, ploidy=2, type="codom")

print(sprintf("Done building bootstrapped dataset %s of %s", bootrep, bootreps))
return(reconvert.genind)
})


```

briefly check out that stuff.

```{r}
print(onethousanddatasets_by_species[7])
print(onethousanddatasets_by_species[[1]])
print(onethousanddatasets_by_species[999])
data.class(onethousanddatasets_by_species[7])
#annoying from a indexing perspective. dataset is a list with 1000 elements, each of them itself is a list, each sublist contains only one thing, a genind file.
print(length(onethousanddatasets_by_species))
print(length(onethousanddatasets_by_species[4]))
data.class(onethousanddatasets_by_species[[7]])
#bootstrapped data tends to have fewer total alleles per dataset, because some not sampled. evens out in grand scheme over many, all information is included. 

filepath <- "1000_geninds_bootstrapped_match_sample_size_by_species_5_5_17.RData" 
save(onethousanddatasets_by_species, file=filepath)


```


Then start popgen calculations.
```{r}
#For time's sake, only run first 100 for a start. 
bootreps <- 100
t0 <- Sys.time()
testresults <- lapply(1:bootreps, function(bootrep){
  ts <- Sys.time()
  boot.genind <- onethousanddatasets_by_species[[bootrep]]
  print(data.class(boot.genind))
  boot.gtypes <- genind2gtypes(boot.genind)
  boot_oT <- overallTest(boot.gtypes, nrep=1000)
  boot_pT <- pairwiseTest(boot.gtypes, nrep=1000)
  tf <- Sys.time()
  print(sprintf("***Done with rep %s of %s, which took %s minutes. %s hours since start.***", bootrep, bootreps, format(as.numeric(tf-ts,units="mins"), digits=3), format(as.numeric(tf-t0,units="hours"), digits=3)))

return(list(boot_oT, boot_pT))
})

filepath <- "100_bootstrapped_popgen_calcs_match_sample_size_by_species_5_5_17.RData" 
save(testresults, file=filepath)



```

Because that was reasonable in time, try the full set, 1000 instead of 100. might finish over weekend.
```{r}
#For time's sake, only run first 100 for a start. 
bootreps <- 1000
t0 <- Sys.time()
testresults <- lapply(1:bootreps, function(bootrep){
  ts <- Sys.time()
  boot.genind <- onethousanddatasets_by_species[[bootrep]]
  print(data.class(boot.genind))
  boot.gtypes <- genind2gtypes(boot.genind)
  boot_oT <- overallTest(boot.gtypes, nrep=1000)
  boot_pT <- pairwiseTest(boot.gtypes, nrep=1000)
  tf <- Sys.time()
  print(sprintf("***Done with rep %s of %s, which took %s minutes. %s hours since start.***", bootrep, bootreps, format(as.numeric(tf-ts,units="mins"), digits=3), format(as.numeric(tf-t0,units="hours"), digits=3)))

return(list(boot_oT, boot_pT))
})

filepath <- "1000_bootstrapped_popgen_calcs_match_sample_size_by_species_5_5_17.RData" 
save(testresults, file=filepath)

```
















