---
title: "Calculating_Bayesian_Deviance_12_runs_20_reps"
author: "Jessie Berta-Thompson"
date: "September 7, 2017"
output: html_document
---
Continue work september 13, 2017

Faubet et al., 2007 & Meirmans 2014 recommend that when running BayesAss you should protext yourself from runs that did not converge (got stuck near boundaries of prior) with the following steps:
1) Run  multiple replicates
2) Calculate Bayesian deviance (a measure of model fit) "to discriminate runs that converged from those that did not"
3) Present only results from the run with the lowest D (rather than averaging replicates)


## Usual warm up clean up shop
```{r}
#clear environment
rm(list=ls())
gc()
getwd()

```

## Calculate Bayesian Deviance for 12 models 20 replicate runs to choose best replicate for analysis
Use formula from script provided by Meirmans in supplemental material (quoted at bottom)
```{r}
#make a list of the directories containing the different model & rep results (BayesAss run with different population definitions)
inputdir <- list.dirs("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/BayesAss/12_runs", recursive=FALSE)
print(sprintf("Found %i model directories.", length(inputdir)))

#loop over the different model directories
Ds_all_models_replicates <- lapply(1:length(inputdir), function(x){
  modelnamesmash = strsplit(inputdir[x], "/")[[1]]
  modelname = modelnamesmash[length(modelnamesmash)]

  #make a list of directories within each model directory - these are the replicates
  repdir <- list.dirs(inputdir[x], recursive = FALSE)
  print(sprintf("Model directory %s contains %i replicate directories.",modelname, length(repdir)))
  
  #make a list containing the probability traces for all replicates within this model
  rep_Ds <- lapply(1:length(repdir), function(y){
    
    # Change this value to the actual burnin used for your MCMC run
    burnin = 3000000
    # Change this value to the actual sampling interval used for your MCMC run
    sampling.interval = 200
    
    #read in data
    whichrep = as.integer(strsplit(repdir[y],"rep")[[1]][2])
    print(sprintf("Working on replicate %i.", whichrep))
    trace=read.table(paste(repdir[y],"/BA3trace.txt",sep=""), header=TRUE)
    
    # Plotting the likelihoods is always a good idea
    pdf(paste("simple_logprob_trace_",modelname,"_rep",whichrep,".pdf", sep = ""))
    plot(trace$State,trace$LogProb, xlab="Iteration", ylab="Log(likelihood)", type="l", lwd=2, col="darkslateblue", font.lab=2)
    # Draw a vertical line to indicate the end of the burnin
    abline(v=burnin, col="grey70", lty=2)
    dev.off()
    
    png(filename=paste("simple_logprob_trace_",modelname,"_",whichrep,".png", sep =""))
    plot(trace$State,trace$LogProb, xlab="Iteration", ylab="Log(likelihood)", type="l", lwd=2, col="darkslateblue", font.lab=2)
    # Draw a vertical line to indicate the end of the burnin
    abline(v=burnin, col="grey70", lty=2)
    rect(xleft = 0, ybottom = -Inf, xright = burnin, ymax = Inf, col= rgb(0.85,0.85,0.85,alpha=0.5))+
    dev.off()
    
    # Calculate the deviance
    range = (trace$State > burnin & trace$State %% sampling.interval == 0)
    D = -2*mean(trace$LogProb[range])

    return(D)
  })
  
  return(rep_Ds)
  
}
)
```



Source Material:

# calculateDeviance.R; Patrick G. Meirmans (2013) Non-convergence in Bayesian estimation of migration rates, Molecular Ecology Resources; Supplementary Material

# This script will calculate the Bayesian Deviance from the output of BayesAss (Wilson & Rannala 2003)
# For more information on the use of the deviance, see Faubet et al. (2007)

# To use this script, run BayesAss version 3 with the -t flag to produce a trace-file
# Then set the working directory of R to the folder with the output from BayesAss

# Change this value to the actual burnin used for your MCMC run
burnin = 1000000
# Change this value to the actual sampling interval used for your MCMC run
sampling.interval = 2000


# Read the data from the trace file
trace=read.table("BA3trace.txt", header=TRUE)

# Plotting the likelihoods is always a good idea
plot(trace$State,trace$LogProb, xlab="State", ylab="LogProb", type="l", lwd=2, col="firebrick3", font.lab=2)
# Draw a vertical line to indicate the end of the burnin
abline(v=burnin, col="grey70", lty=2)

# Calculate the deviance
range = (trace$State > burnin & trace$State %% sampling.interval == 0)
D = -2*mean(trace$LogProb[range])

# Print the result to the console
print(D)



