---
title: "Examining_results_of_strataG_for_1000X_bootstraps_by site_Fst_calcs_only"
author: "Jessie Berta-Thompson"
date: "May 8, 2017"
output: html_document
---


*************************************************************************************
## Set up
Clear history and make sure you're in the right working directory for shorter file writes. 
load in strataG results.
```{r}
rm(list=ls())
print("Where are we? Where will new files without specified full path be created?")
getwd()

#Get results file of interest
load("1000_bootstrapped_popgen_calcs_match_sample_size_by_location_onlyFst_pairwise_5_5_17.RData")

#load in results for straight calculations (not bootstrapped) too.
load("full_data_strataG_results_5_4_17.RData")

```

Look at results structure to understand how to access things
```{r}
print("how big is the saved results object and what is it?")
print(length(testresults))
print(data.class(testresults))
#list of 1000 replicate results.

print("how big is one element of results list?")
print(length(testresults[1]))
print(length(testresults[[1]])) #have to index with double brackets.

#What's in each replicate's sublist?
print(length(testresults[[1]][[1]])) #also benefits from double bracketting.
print(length(testresults[[1]][[2]]))

print(testresults[[1]][[1]]) #strataG overall test result
print(testresults[[1]][[2]]) #strataG pairwise Fst result

#Results for one replicate dataset
print(testresults[[1]][[2]]$result$pair.label) #list of labels for pairwise comparisons
print(testresults[[1]][[2]]$result$Fst) #Fst values for pairwise comparisons between sites
print(testresults[[1]][[2]]$result$Fst.p.val) #p values for permutation test of Fst values (diff from Fst = 0, panmictic)

#how many pairwise comparisons were there for 15 populations?
print(length(testresults[[1]][[2]]$result$Fst))
#105.

#Order same for all?
print(testresults[[1]][[2]]$result$pair.label)
print(testresults[[5]][[2]]$result$pair.label)
print(testresults[[736]][[2]]$result$pair.label)
print(testresults[[820]][[2]]$result$pair.label)
print(testresults[[999]][[2]]$result$pair.label)

testresults[[820]][[2]]$result$pair.label == testresults[[999]][[2]]$result$pair.label



```

*************************************************************************************
## Parse test results object to get FST for 105 pairwise comparisons
```{r}
nboot <- length(testresults)
ncompare <- length(testresults[[1]][[2]]$result$Fst)
print(sprintf("This dataset contains %s bootstrap replicates, each with %s pairwise comparisons between sites.", nboot, ncompare))

#Order same for all, just grab that once.
labels <- testresults[[1]][[2]]$result$pair.label

#Grab Fst values in list of lists: loop over replicates
list_of_vectors <- lapply(1:nboot, function(n) {
  #grab vector of Fst results for 105 comparisons.
  fsts105 <- testresults[[n]][[2]]$result$Fst
  return(fsts105)
})

#turn that into a dataframe
fsts.df <- do.call("rbind", list_of_vectors)
print(data.class(fsts.df))
print(dim(fsts.df))
#rows are replicates, columns are comparisons.
colnames(fsts.df) <- labels #label columns with pairwise comparisons
print(fsts.df[1:10,1:5]) 
rownames(fsts.df) <- c(1:1000)
print(fsts.df[1:10,1:5]) 

colMeans(fsts.df)
means <- apply(fsts.df, 2, mean)
standarddeviations <- apply(fsts.df, 2, sd)
```
Quick version, warm up
```{r}
# Violin Plots
#install.packages('vioplot')
library(vioplot)
#jpeg( "violin_1000xbootstraps_pairwiseFst_by_species.jpg", width = 600, height = 600 )
pdf( "violin_1000xbootstraps_pairwiseFst_by_location.pdf", width = 12, height = 5 )

vioplot(fsts.df[,1],fsts.df[,2],fsts.df[,3],fsts.df[,4],fsts.df[,5], names=labels[1:5], 
   col="purple", ylim = c(0, 1.0))

title(main="1000x bootstrapped Fst values pairwise by site ", ylab="Fst")

```

Work on reformatting for ggplot (melted data, with all the extra info that we want)
```{r}
library(reshape2)
melted <- reshape2::melt(fsts.df, value.name = 'Fst', varnames = c('Replicate', 'Comparison'))
#print(data.class(melted))
#print(dim(melted))
#print(melted[1:5,])

list_of_comparisons <- as.character(melted$Comparison)
#data.class(list_of_comparisons)
#length(list_of_comparisons)

first_thing<- unlist(lapply(list_of_comparisons, function(compstring) {
  halves <- strsplit(compstring, "v. ")
  messy_first <- halves[[1]][[1]]
  tidying <- strsplit(messy_first, " ")
  tidied <- tidying[[1]][[1]][[1]]
}))

second_thing<- unlist(lapply(list_of_comparisons, function(compstring) {
  halves <- strsplit(compstring, "v. ")
  messy_second <- halves[[1]][[2]]
  tidying <- strsplit(messy_second, " ")
  tidied <- tidying[[1]][[1]][[1]]
}))

print(first_thing[1:5])

melted$Site1 <- first_thing
melted$Site2 <- second_thing
print(melted[1:5,])




scullyi <- as.character(c("HN", "HS", "F"))
formosula <- as.character(c("CG", "D", "P","RN", "WS", "WA", "B", "V", "R", "AN", "AS"))
gina_glenneae <- as.character(c("T"))


sitelist1 <- as.character(melted$Site1)
species1 <- unlist(lapply(sitelist1, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

print(length(species1))
print(data.class(species1))
print(species1[1:5])

sitelist2 <- as.character(melted$Site2)
species2 <- unlist(lapply(sitelist2, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

print(length(species2))
print(data.class(species2))
print(species2[1990:2005])

melted$Species1 <- species1

melted$Species2 <- species2


print(melted[1995:2005,])



```

#County another way to think about these
formosula  = Jackson
gina-glenneae = Grand
scullyi = Larimer

other ways of slicing data 

```{r}

same_or_mixed <- unlist(lapply(1:length(melted$Species1), function(i) {
  if ( melted$Species1[i] == melted$Species2[i] ) {
    result <- "same county"
  }
  else {
    result <- "different county"
  }
  return(result)
}))

species_pair <- unlist(lapply(1:length(melted$Species1), function(i) {
  pair <- paste(melted$Species1[i], "vs.", melted$Species2[i]) 
}))

melted$samenot <- factor(same_or_mixed)
melted$county_compare <- factor(species_pair, levels = c("formosula vs. gina-glenneae", "formosula vs. formosula", "scullyi vs. formosula", "scullyi vs. scullyi", "scullyi vs. gina-glenneae"))

        
  


```



Make a plot! 
```{r}
library(ggplot2)
pdf(file = "1000XbootstrappedFstbysite_violin_color_by_county_same_or_diff.pdf", width = 12, height = 5)
ggplot(melted, aes(x = melted$Comparison, y = melted$Fst, fill = melted$samenot)) + 
  geom_violin() +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("1000x Bootstrapped Fst values for all pairwise comparisons between 15 sites")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")
```

Make a plot! color by species pair
```{r}
library(ggplot2)
pdf(file = "1000XbootstrappedFstbysite_violin_color_by_county_comparison.pdf", width = 12, height = 5)
ggplot(melted, aes(x = melted$Comparison, y = melted$Fst, fill = melted$county_compare)) + 
  geom_violin() +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("1000x Bootstrapped Fst values for all pairwise comparisons between 15 sites")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_fill_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF6000" ))
```

Make a plot! color by species pair
sort by species pair
```{r}
library(ggplot2)
sorted <- melted[order(melted$county_compare),]
head(sorted)
shorter<-unique(sorted$Comparison)
data.class(shorter)
shorter <- as.character(shorter)
sorted$Comparison <- factor(sorted$Comparison, levels = shorter)



pdf(file = "1000XbootstrappedFstbysite_violin_color_by_county_comparison_sorted_by_pair.pdf", width = 12, height = 5)
ggplot(sorted, aes(x = sorted$Comparison, y = sorted$Fst, fill = sorted$county_compare)) + 
  geom_violin() +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("1000x Bootstrapped Fst values for all pairwise comparisons between 15 sites")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_fill_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF6000" ))
```





sort by mean value: get mean value with names, sort, change factor order. 
```{r}
means <- apply(fsts.df, 2, mean)
sorted_means <- sort(means)
sorted_names <- names(sorted_means)

#change factor levels order
sorted_by_mean <- sorted
sorted_by_mean$Comparison <- factor(sorted_by_mean$Comparison, levels = sorted_names)


pdf(file = "1000XbootstrappedFstbysite_violin_color_by_county_comparison_sorted_by_value.pdf", width = 12, height = 5)
ggplot(sorted_by_mean, aes(x = sorted_by_mean$Comparison, y = sorted_by_mean$Fst, fill = sorted_by_mean$county_compare)) + 
  geom_violin() +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("1000x Bootstrapped Fst values for all pairwise comparisons between 15 sites")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_fill_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF6000" ))
```


sort by mean value: add actual calculations (not bootstrap) 
```{r}

#extract from their datastructure
fst <- as.numeric(real_by_loc_pT$result$Fst)
fst_names <- real_by_loc_pT$result$pair.label

realfst.df <- data.frame(Pair=fst_names, Fst = fst)
head(realfst.df)

long_real <- unlist(lapply(1:length(sorted_by_mean$Comparison), function(i) {
  comp <- sorted_by_mean$Comparison[i]
  #print(comp)
  index <- which(realfst.df$Pair == comp)
  #print(realfst.df$Pair[index])
  #print(realfst.df$Fst[index])
  return(realfst.df$Fst[index])
}))

sorted_by_mean$actual_Fst <- long_real
print(sorted_by_mean$actual_Fst[1:5])
data.class(sorted_by_mean$actual_Fst)

```
```{r}
pdf(file = "1000XbootstrappedFstbysite_violin_color_by_county_comparison_sorted_by_value_with_real.pdf", width = 12, height = 5)

ggplot() + 
  geom_violin(data=sorted_by_mean, mapping=aes(x = sorted_by_mean$Comparison, y = sorted_by_mean$Fst, fill = sorted_by_mean$county_compare)) +
  geom_point(data=sorted_by_mean, mapping=aes(x = sorted_by_mean$Comparison, y=sorted_by_mean$actual_Fst))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("1000x Bootstrapped Fst values for all pairwise comparisons between 15 sites")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_fill_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF6000" ))
```








Make a plot! color by species pair
sort by species pair
add real data
```{r}
long_real2 <- unlist(lapply(1:length(sorted$Comparison), function(i) {
  comp <- sorted$Comparison[i]
  #print(comp)
  index <- which(realfst.df$Pair == comp)
  #print(realfst.df$Pair[index])
  #print(realfst.df$Fst[index])
  return(realfst.df$Fst[index])
}))

sorted$actual_Fst <- long_real2
print(sorted$actual_Fst[1:5])
```


```{r}
pdf(file = "1000XbootstrappedFstbysite_violin_color_by_county_comparison_sorted_by_pair_with_real.pdf", width = 12, height = 5)

ggplot() + 
  geom_violin(data=sorted, mapping=aes(x = sorted$Comparison, y = sorted$Fst, fill = sorted$county_compare)) +
  geom_point(data=sorted, mapping=aes(x = sorted$Comparison, y=sorted$actual_Fst))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("1000x Bootstrapped Fst values for all pairwise comparisons between 15 sites")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_fill_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF6000" ))
```










**************************************************************************************************************************
## Plot just real data using schemes developed here

# Create a dataframe 
with real data, melt it, and build up collection of plot-relevant factors, with meaningfully ordered and sorted levels.

```{r}
#Extract data and labels from strataG result object
fst <- as.numeric(real_by_loc_pT$result$Fst)
pair_label <- real_by_loc_pT$result$pair.label

#Bundle those into a dataframe
fst.df <- data.frame(Pair=pair_label, Fst = fst)
head(fst.df)




#Extract the two sites from the label as separate quantities and add to dataframe as additional factors
first_site<- unlist(lapply(pair_label, function(compstring) {
  halves <- strsplit(compstring, "v. ")
  messy_first <- halves[[1]][[1]]
  tidying <- strsplit(messy_first, " ")
  tidied <- tidying[[1]][[1]][[1]]
}))

second_site<- unlist(lapply(pair_label, function(compstring) {
  halves <- strsplit(compstring, "v. ")
  messy_second <- halves[[1]][[2]]
  tidying <- strsplit(messy_second, " ")
  tidied <- tidying[[1]][[1]][[1]]
}))

fst.df$Site1 <- factor(first_site)
fst.df$Site2 <- factor(second_site)


# Add Species designations for first and second of comparison pairs.
scullyi <- as.character(c("HN", "HS", "F"))
formosula <- as.character(c("CG", "D", "P","RN", "WS", "WA", "B", "V", "R", "AN", "AS"))
gina_glenneae <- as.character(c("T"))


species1 <- unlist(lapply(first_site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

species2 <- unlist(lapply(second_site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

fst.df$Species1 <- factor(species1)
fst.df$Species2 <- factor(species2)

#Add Counties (equivalent to species; but perhaps we'd want to label things geographically)
county1 <- unlist(lapply(species1, function(sp) {
  if (sp == "scullyi") {
    co <- "Larimer"
  }
  else if (sp == "formosula") {
    co <- "Jackson"
  }
  else if (sp == "gina-glenneae") {
    co <- "Grand"
  }
  else {
    print("uh-oh. mapping fail.")
  }
}))

county2 <- unlist(lapply(species2, function(sp) {
  if (sp == "scullyi") {
    co <- "Larimer"
  }
  else if (sp == "formosula") {
    co <- "Jackson"
  }
  else if (sp == "gina-glenneae") {
    co <- "Grand"
  }
  else {
    print("uh-oh. mapping fail.")
  }
}))

fst.df$County1 <- factor(county1)
fst.df$County2 <- factor(county2)


#Add combinations of species and counties
species_pair <- unlist(lapply(1:length(species1), function(i) {
  pair <- paste(species1[i], "vs.", species2[i]) 
}))

fst.df$SpeciesComparison <- factor(species_pair, levels = c("formosula vs. gina-glenneae", "formosula vs. formosula", "scullyi vs. formosula", "scullyi vs. scullyi", "scullyi vs. gina-glenneae"))

counties_pair <- unlist(lapply(1:length(species1), function(i) {
  pair <- paste(county1[i], "vs.", county2[i]) 
}))

fst.df$CountyComparison <- factor(counties_pair, levels = c("Jackson vs. Grand","Jackson vs. Jackson","Larimer vs. Jackson", "Larimer vs. Larimer", "Larimer vs. Grand"))


melted$samenot <- factor(same_or_mixed)
melted$county_compare <- factor(species_pair, levels = c("formosula vs. gina-glenneae", "formosula vs. formosula", "scullyi vs. formosula", "scullyi vs. scullyi", "scullyi vs. gina-glenneae"))

        
#Add one more, sorting by whether pair in a comparison is from the same county/species or different.
same_or_mixed <- unlist(lapply(1:length(species1), function(i) {
  if ( species1[i] == species2[i] ) {
    result <- "same species" #if you prefer county/species language, change these strings
  }
  else {
    result <- "different species"
  }
  return(result)
}))

fst.df$SameOrDiff <- factor(same_or_mixed)

print(fst.df)
```

```{r}
#So organized! Time to plot! In different orders! Other stats too, not just Fst - functionitize and repeat when done. 
library(ggplot2)
pdf(file = "real_data_pairwise_Fst_by_comparison_species_5_9_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=fst.df, mapping=aes(x = fst.df$Pair, y=fst.df$Fst,  color = fst.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF6000" ))
  

```









#others
plot boots sorted by real values
plot just real values with color scheme, add element on p value (alpha)
sort by real value: get mean value with names, sort, change factor order. 
Plot Fsts from real data in same color scheme/styles.
Also: Sort by species comparison, then by value.
Also: Sort by value, color by within or between species.




