---
title: "Building MetaPopGen scripts for Phacelia: formatting output for popgen calculations"
author: "Jessie Berta-Thompson"
date: "March 1, 2017"
output: html_document
---

Purpose of this version:
Work on converting output into more pop-gen calculation friendly form (genind object)

Working from Nederland: 
10 AM start - 10:30

Thoughts for future work as they come up:

 - Eventually will continue exploring demographic stochasticity of bad years (make a version that's stochastic across replicates), but for now behavior of that one was odd - just work with a flatter demography while building up tools. Too big of swings in current. 

 - Population structure set by Kappa and survival rates. build a little tool to pick those values in order to create realistic population structure (current one a little flower-heavy)
 
- write generic code for making print tables (e.g. tabs, padding, like python format)

**************************************************************************************
Set up: clear history and make sure you're in a good working directory (files will be generated), load program.
```{r}
rm(list=ls())
getwd()
library(MetaPopGen)
```


Generate test data by running example case with some similarity to Phacelia system
# Part 1: Creating the initial dataset
```{r}
n <- 3		# Number of demes - 3 populations
z <- 2		# first year rosettes, second year flowering plants.
l <- 10		# Number of alleles -  observations average 8.5/locus; model 10
m <- l*(l+1)/2		# Number of genotypes = diploid combinations of alleles l(l+1)/2
tmax <- 1000		# Number of years of simulation

	# Initialize matrices that will hold survival and reproduction rates
  # using the correct sizes for each dimension
  # Dimensions = Genotypes x Demes x Ages x Time
		sigma <- array(NA,dim=c(m,n,z,tmax)) #survival 
		phi_M <- array(0,dim=c(m,n,z,tmax)) #male fecundity
		phi_F <- array(0,dim=c(m,n,z,tmax)) #female fecundity
  
		# Important to track order of events for interpreting this: 
	# 1) survival - multiply input  by survival, removes some rosettes and flowering plants.
	# 2) reproduction - flowering plants produce seed
	# 3) recruitment - some seed becomes rosettes, rosettes become flowers, flowers  removed
	# N1 is before you start any thing, N2 is after one cycle of this.
		
		# This assigns values to the variables
		sigma[,,1,] <- 0.3 #survival is 30% for rosettes 
		sigma[,,2,] <- 0.9 #survival is 90% for flowering
		phi_M[,,2,] <- 10000 #flowering males make 10,0000 pollen grains each
		phi_F[,,2,] <- 1000 #flowering females make 500 ovules each ~seed count

	# Seed Dispersal matrix
	# seeds move around before sprouting into rosettes in recruitment phase
	m.m <- 0.1 # migration - total of 10% come from elsewhere, evenly split between other pops
	perpop <- m.m/(n-1)
	delta <- matrix(rep(perpop,n^2),nrow=n,ncol=n)
	diag(delta)<-rep((1-m.m),n)
	
	# Initial population structure:
	# flat - all genotypes, all demes have the same number of individuals; different number of rosettes and flowering plants.
	# Initialize array for initial number of individuals for [genotypes * demes * ages]
		N1 <- array(NA,dim=c(m,n,z))		
		# different starting n individuals for different age classes
		# entering values that are per population/deme, per genotype
		nrosettes = 100 #lots of rosettes
		nflowers = 10 #some flowering
    N1[,,1] <- nrosettes #fill in value across all genotypes, across all demes, age class 1
    N1[,,2] <- nflowers #fill in value across all genotypes, across all demes, age class 2
    
    
	# Mutation matrix - uses max alleles at locus to model k-alleles mutation
    overall_mu <- 1e-5 #likelihood of an allele mutating to any another allele
    not_mutating <- 1-overall_mu
    per_allele_mu <- overall_mu/(l-1) #there are l-1 nonself alleles to mutate to, each with equal likelihood 
    #** could change this to model stepwise - 50% chance of moving up or down 1,with careful handling of edges.
		mu <- array(per_allele_mu,dim=c(l,l)) # fill in array with per_allele_mu
		#then make the diagonals the not_mutating rate
		for (from_allele in 1:l){
		  for (to_allele in 1:l){
		    if (from_allele==to_allele) {
		      mu[from_allele, to_allele] <- not_mutating
		    }
		  }
		}
#see how it came out
print("print mutation rate matrix, between alleles")
print(mu)

		
	# Maximum recruitment per deme: Juvenile Carrying Capacity
  # After reproducing, how many will make it into next year's starting Age 1
  # Flat kappas
	kappa0 <- array(5000,c(n,tmax)) #at most 5000 seedlings allowed per population.
  # excess of seed in system means that kappa usually reached, randomness near asymptote, not limited by seed
		
	# Add good and bad years: this part needs some work, once evaluative tools up and running
	#	kappas <- c(500, 5000) #number of seedlings in a bad year, number of seedlings in a good year
	#	probs <- c(0.2, 0.8) # probability of bad year, good year
	#	qualities <- vector(length=tmax)
	#	kappa0 <- array(NA, c(n, tmax))
	#	for (tp in 1:tmax) {
	#	  kappa0[,tp] <- sample(kappas, 1, prob=probs)
	#	}
		#this will be the same bad years across all replicates - might be more useful later to put randomness across reps.
		
# Save parameter set
# usually something like
# filepath <- "Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/test_V6_ph_test_2_28_17.RData" 

# today:
filepath <- "/Users/jessie/Documents/botanic/MetaPopGen_coding_Feb_2017/test_V7_phacelia_3_1_17_parameters.RData"
save(sigma,N1,phi_M,phi_F,mu,delta,kappa0,tmax, file=filepath)


```

See how N1 came out - is the number of individuals set up to be different across age classes?
```{r}
print("N1 Dimensions = genotype * deme * age class")
print(dim(N1)) # 6 genotypes x 4 demes x 3 age classes
print("Genotype 1, deme 1, age 1 and 2")
print(N1[1,1,]) #look at 1 genotype, 1 deme, all age classes
print("Genotype 50, deme 3, age 1 and 2")
print(N1[50,3,]) #look at 1 genotype, 1 deme, all age classes
print(sprintf("%s rosettes per deme, %s flowering plants per deme to start", sum(N1[,1,1]), sum(N1[,1,2])))

```

#############################################################################################
# Perform the simulation
#############################################################################################

	
```{r}
#Optional: load saved workspace (starting from here)

#library(MetaPopGen)
#rm(list=ls())
#filepath <- "Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/test_V5_ph_test_2_23_17.RData"
#load(filepath)

#Alternative: working from variables saved above.

	# Simulation
		nrepl <- 10

  # Create a place to store Ns as you go
		Ns <- array(NA, dim=c(nrepl, m, n, z, tmax))
		
	# Create an array to store fst for each rep, each time point
		fst <- array(NA,dim=c(nrepl,tmax))
	
	# Run the simulations
	for (i in 1 : nrepl) {
	  
	  cat("Replicate",i,"\n") # Just to print on screen the advancement of the simulation
	  
	  N<-sim.metapopgen.monoecious(
		input.type="array",
		N1 = N1, sigma = sigma,
		phi_M = phi_M, phi_F = phi_F,
		mu = mu, delta = delta, kappa0 = kappa0,
		T_max = tmax,
		save.res = F)
		
		flush.console() #ensures that display of output in the console is current, even if output buffering is on.
		
		#store the N, after calculating a replicate, to an array 
		Ns[i,,,,] <- N
		
		# Calculate global Fst for replicate across timepoints
		print("Calculating fst...")
		for (t in 1 : tmax){
			fst[i,t] <- fst.global.monoecious(N,t)
		}
	}

filepath <- "/Users/jessie/Documents/botanic/MetaPopGen_coding_Feb_2017/test_V7_phacelia_3_1_17_results.RData"
save(Ns,fst,sigma,N1,phi_M,phi_F,mu,delta,kappa0,tmax,nrepl,file=filepath)
#load(filepath) #to load saved results. 
```		

Examine output.
```{r}
#Ns has dimensions [replicates * genotypes * demes * ages * times]
print("dimensions")
print(c("replicates", "genotypes", "demes", "ages", "times"))
print(dim(Ns))
print(data.class(Ns))
```

Exploring subsampling methods
```{r}
#How many years between taking samples for analysis?
interval <- 10 # sample every n years

#build a list of sampling generations based on this interval.
samplemid <- seq.int(interval, tmax-interval, interval)
samplegens <- c(1, samplemid, tmax)
print("Sampling times")
print(samplegens)

```

Print statements too big for full dataset, but looping patterns and calculations might be useful.
```{r}
#summarize population for a snapshot
for (d in 1:n) { 
  #for a subset of times:
  for (tp in samplegens){
  year = Ns[1,,d,,tp] #everything for time tp and deme d replicate 1 (all genotypes and ages)
  print(sprintf('Year %s, deme %s, average number of individuals per genotype for 2 age classes',tp, d))
  print(apply(year, 2, mean)) #calculate the mean individuals across genotypes - dimension 2 of genotype vs age matrix
}
}

#summarize 
for (d in 1:n) { 
  #for a subset of times:
  for (tp in samplegens){
  year = N[,d,,tp] #everything for time tp and deme d
  print(sprintf('Year %s, deme %s',tp, d))
  print("total individuals")
  print(apply(year, 2, sum))
  }
}

#One deme, all years
  for (tp in 1:tmax){
  year = N[,d,,tp] #everything for time tp and deme d
  print(sprintf('Year %s, deme %s',tp, d))
  print("total individuals")
  print(apply(year, 2, sum))
}

```

Exploring slices of counts with plots
```{r}
#plot
#one rep, one genotype, one deme, one age class, all times
one_slice = Ns[1,1,1,1,]
par(mar=c(5,10,5,5)) #bottom, left, top, right; default c(5, 4, 4, 2) + 0.1.
plot(1:tmax,one_slice, ylab="Number of individuals for \n one replicate, one genotype, one deme, \n just rosettes", xlab="Years")
lines(1:tmax,one_slice)


#sum over genotypes - rosettes
par(mar=c(5,10,5,5)) #bottom, left, top, right; default c(5, 4, 4, 2) + 0.1.
summed_genotypes <- lapply(1:tmax, function(t) {
  all_genotypes <- Ns[1,,1,1,t]
  sum(all_genotypes)}
)
plot(1:tmax, summed_genotypes, xlab="years", ylab="individuals of all genotypes,\n one replicate, one deme, rosettes only")
lines(1:tmax, summed_genotypes)


#sum over genotypes, just flowering
summed_genotypes <- lapply(1:tmax, function(t) {
  all_genotypes <- Ns[1,,1,2,t]
  sum(all_genotypes)}
)
par(mar=c(5,10,5,5)) #bottom, left, top, right; default c(5, 4, 4, 2) + 0.1.
plot(1:tmax, summed_genotypes, xlab="years", ylab="individuals of all genotypes,\n one replicate, one deme, flowering only")
lines(1:tmax, summed_genotypes)

```











Kappa: after everything else, perhaps at same time as Age class 1 & 2 remaining after survival moves up a class to 2&3 (and class3 disappears) , newly created class 1 individuals fill in empty class 1, sampling from individuals produced during reproduction to level allowed by kappa. from paper: "Number of recruits tends assymptotically to" kappa. Probably sets of stable conditions if number of offspring exceeds kappa. Hurts noisiness of data (but still sampling a binomial)


******************************************************************
Next Question:
How to get alleles from genotypes?
Mutation happens at level of l (# alleles), and alleles are combined to create genotypes (pairs of alleles), but the data is in the form of just genotype, not clear which genotypes share alleles, which are homo/heterzygotes. 

Documentation of example "FourPopMonoecious" includes lines defining order of genotypes.

barplot(N1[,,1],names.arg=c("1","2","3","4"),xlab="Deme",ylab="Number of age-1 individuals",col=colours)
lg.txt <- c("A1A1", "A1A2", "A1A3", "A1A4", "A2A2", "A2A3", "A2A4", "A3A3", "A3A4", "A4A4")
legend(-1,1500,lg.txt,fill=colours,ncol=5,bty="n")

barplot(N[,,2,dim(N)[4]],names.arg=c("1","2","3","4"),xlab="Deme",ylab="Number of age-2 individuals",col=colours)
lg.txt <- c("A1A1", "A1A2", "A1A3", "A1A4", "A2A2", "A2A3", "A2A4", "A3A3", "A3A4", "A4A4")
legend(0.2,600,lg.txt,fill=colours,ncol=5,bty="n")

And even easier to understand, elsewhere in this example they state it:

"If there are two alleles, say A1 and A2, there will be three genotypes; the order in which they are stored in N1 is as follows: A1A1, A1A2 and A2A2. If there are three alleles (A1, A2 and A3), the order is: A1A1, A1A2, A1A3, A2A2, A2A3, A3A3. With four alleles, the order is: A1A1, A1A2, A1A3, A1A4, A2A2, A2A3, A2A4, A3A3, A3A4, A4A4. So with four alleles N[6,1,3] gives the number of A2A4 individuals in deme 1 that are of age 3."

******************************************************************

******************************************************************

******************************************************************
Next Question:
How should i assign life cycle stages to these yearly age classes?

Age 1 = sprouts (mini rosettes, recruited seeds)
Age 2 = flowering plants (reproductive)

-----T = 1 starting conditions

Survive:
some sprouts don't survive (this means they won't move up to flower)
some flowering plants don't survive (this means they won't produce seed)

Reproduce:
Only flowering plant class 2 reproduces, but they make a lot of pollen and many ovules (ovules sets amount of seed)

Recruit:
A density dependent amount of seed actually sprouts (can't exceed carrying capacity)
Last year's flowers fall off the map
Last year's sprouts move up to flowering.

-----T = 2 make another count at this point

******************************************************************

******************************************************************

******************************************************************
Next Question:
How can I calculate popgen stats from the output of this program?

Built-in tools:

##genotype.index(l)

Give the number of alleles, it shows you the genotype number (index along genotype axis of N matrix) for each combination of alleles. useful for translating between data matrix and actual allelic data.

```{r}
genotype_indices = lapply(2:10, genotype.index)

for (i in 1:length(genotype_indices)){
  print(genotype_indices[i])
}

```


### create.dispersal.matrix(x,a,longlat = True)
There is an option to input a matrix of spatial data x = lat/lon matrix; row denotes points, first column longitude, second column latitude (where are your demes), output distance-dependent migation rates. 

##freq.all(N)
input number of individuals of each genotype (N[,1,1,1])
```{r}
#for deme 1, age 1
print(freq.all(N[,1,1,1])) #starting frequences of A1, A2 and A3
print(freq.all(N[,1,1,5])) #ending frequencies

#slicing question, can I get a few at once?
print(freq.all(N[,,,5])) #all demes and all age classes
#no! this makes no sense - more frequencies listed than alleles. 
#will require some slicing and squishing in different ways. 
#how many total genotypes?
print(dim(N))
```


## fst.global.monoecious
## fst.global.dioecious
Compute Fst between two groups of individuals (dioecious or mono)

fst.pairwise.monoecious(N,i,j,xi,xj,ti,tj)
fst.pairwise.monoecious(folder.name,i,j,xi,xj,ti,tj)

N	
A 4-D array (ex: N[k,i,x,t]) of length c(m,n,z,T_max) giving the number of individuals of genotype k, age x in deme i at time t. Typically, the output of the function sim.metapopgen.monoecious
folder.name	
Character string giving the name of the folder where the results of sim.metapopgen.monoecious are saved, when this function is used with the argument save.res set to TRUE.
i, xi, ti	
Identifiers of the first group of individuals: it specifies the individuals of deme i, age xi at time ti
i, xi, ti	
Identifiers of the second group of individuals: as above.



```{r}
print(dim(N))
fst.pairwise.monoecious(N,1,2,2,2,5,5) #compare time 5, age 2, Deme 1 vs 2.
tmax = 5
# Temporal evolution of fst
fst <- array(NA,dim=c(3,5))
for (t in 1 : tmax) {
for (deme in 1 : 3){
fst[deme,t] <- fst.pairwise.monoecious(N,1,(deme+1),1,1,t,t)
}
}
plot(c(1:tmax),fst[1,],type="l",col="blue",xlab="Year",ylab="Fst")
matplot(c(1:tmax),fst[2,],type="l",col="darkgreen",add=T)
matplot(c(1:tmax),fst[3,],type="l",col="red",add=T)
legend(30,1,c("1 vs 2","1 vs 3","1 vs 4"), col = c("blue", "darkgreen", "red"),lty=1)

```

So, to get more, we'll need to make a function that converts this object into genind

Individual = row
Each row assigned to a population
plus the data column, the locus, with allele assignments

1-2

dig up old code for making genind, figure out how to extract right kinds of data, and modify.

#for each timepoint there will be one dataset.
N = genotypes * demes * age * time

```{r}
#results <- vector("list", n) #make list of defined size
demecount = 4
tmax = 5
for (i in 1:tmax) {#each timepoint a different popgen dataset
  dataset <- N[,,2,i] #within each deme, only look at flowering individuals
  #print(sprintf("For time %s, flowering plants only, genotypes vs. deme", i))
  #print(dataset)
  for (j in 1:demecount){ #go through deme list
    print(sprintf("For time %s deme %s, flowering plant genotype counts", i,j))
    print(N[,j,2,i])
    genotype_vector <- N[,j,2,i]
    print("Length of genotype vector")
    print(length(genotype_vector))
    print("Sum of genotype vector individuals counts = number of flowering plants in deme")
    print(sum(genotype_vector))
    print(data.class(genotype_vector))
    genotypes = 6
    deme_list <- list()
    ind_count = 0
    for (g in 1:genotypes) {
      geno_count <- N[g,j,2,i] #this many of this type
      if(geno_count != 0) {
        for (ind in 1:geno_count) {
          deme_list[length(deme_list)+1] <- g
          ind_count = ind_count + 1
      }
        }
    }
    print(ind_count)
    print(length(deme_list))

      
    }
  }
  
#Next step sample 30 individuals per deme, then in an outer loop be building for each timepoint (data vector) (deme label vector) in as close to format for genind as possible. build these lists into a bigger list convert to 1-1 notation.
 

  
  
```


#Pop stats with sampling from Rmetasim work
```{r}
library(adegenet) # multivariate genetic data analysis
library(strataG)  # summaries and population structure analyses of genetic data

sampled_genind_lol<- lapply(1:nreps, function(r) {# loop over replicates
  sampled_genind_list_over_time<-lapply(1:length(replist[[r]][[2]]), function(t){# loop over matrices 
    print(sprintf("Working in replicate %s, time %s", r, t))
    rland.df <- data.frame(replist[[r]][[2]][t]) #convert one individuals matrix to dataframe format
    #reformat allele data (pairs by locus)
    alleles <- seq(7,length(rland.df)-1,2) # in rland$individuals, alleles start at 7, go to end, pair at each locus side by side
    rland.2genind <- sapply(alleles, function(x){
      paste(rland.df[,x], rland.df[,x+1], sep = "-")}) # Combine alleles per locus with '-' for genind
    #convert to data frame to match the other one.
    rland.2genind <- data.frame(rland.2genind)
    ### For debugging: 
    print("Dimensions of full rland.df (rxc):")
    print(dim(rland.df))
    print("Dimensions of full rland.2genind (rxc):")    
    print(dim(rland.2genind))

    #label with populations [subpopulations 0, 1, 2, 3 = pop1, 4, 5, 6, 7 = pop2, 8, 9, 10, 11 = pop3]
    rland.df$pop <- "pop3"
    rland.df$pop[rland.df$X1 < 8] <- "pop2"
    rland.df$pop[rland.df$X1 < 4] <- "pop1"
    
    #sample just flowering plants from each population: 3, 7, and 11s for 3 populations with 4 age classes in each.
    flowers <- c(3,7,11)
    #sample at random, n individuals per population.
    n <- 30 #number of samples to take. will sample exactly that many - doesn't bother if fewer than 30 flowers.
    sample_rows <- lapply(flowers, function(x){ #loop through flower pops to create sample list.
      n_pop <- length(rland.df$X1[rland.df$X1 == x]) #count the number of instances of given population x. X1 contains pop id (3, 7, 11)
      print(sprintf("population %s contains %s plants", x, n_pop))
      if (n_pop>=n) {
        print(sprintf("enough plants in population %s to sample target %s individuals!",x,n ))
        rowstotake <- sample(which(rland.df$X1 == x),n) #sample at random 30 items from the indices of rows in correct pop
      }
      else{
        print(sprintf("not enough plants in population %s to sample target %s individuals: returning NULL",x,n ))
        rowstotake <- NULL
      }
      return(rowstotake)
      })    
     print(sample_rows) #this is a list of lists
    rows <- do.call(c,sample_rows) #this collapses list of lists into a single vector.
    #this doesn't have special case handling when all three are below 30 - ignore for now, but could be a problem.
    print(rows)
    #apply sampling to two data frames
    samp.rland.2genind <- rland.2genind[rows,] # sample genind-style data (just alleles) for the sample rows, all columns
    samp.rland.df <- rland.df[rows,]

    ### For debugging: 
    print("Dimensions of samp.rland.2genind (rxc):")
    print(dim(samp.rland.2genind))
    print("Dimensions of samp.rland.df (rxc):")    
    print(dim(samp.rland.df))
    
    #convert to genind for downstream analyses
    samp.rland.genind <- df2genind(samp.rland.2genind, pop = samp.rland.df$pop, type = "codom", sep="-") #error occurs here - size mismatch
       return(samp.rland.genind)})
    return(sampled_genind_list_over_time)})
```


#bit of code that might be useful, principle of looping through data structure
```{r}
		
		for (deme in 1 : n){ #loop over demes
		  for (genotype in 1 : m) { #loop over genotypes
				for (age in 1 : z) { #loop over age classes
          if (age == 1) {
					  N1[genotype,deme,age] <- 30 
          }
					else if (age == 2) {
					  N1[genotype,deme,age] <- 20
					}
					}
				}
			}

```



#snippet from examples comparing modeled and expected Fst
# we want to do this, but example was a little too specific to case.
```{r}
# Mean and standard deviation of FST
mean(fst[,3:tmax]) # because reached stable Fst after ca n generations?
sd(fst[,3:tmax])

# Expected value for FST
Ne <- 6*(30 + 15 + 7.5) #each deme starting size 10 + 20 + 30, Kappa 180 limit to incoming.
#per genotype (of 6) 30 - 15 - 7.5 
m.m <- 0
fst.exp <- 1/(1 + 4 * Ne * (m.m))
```



