---
title: "Running MetaPopGen v3"
author: "Jessie Berta-Thompson"
date: "April 5, 2017"
output: html_document
---
Purpose of this version, immediate tasks:
Parallelize: 
Desired model (below) was causing problems, taking too long on the computer needed for other things. 
To make it go faster, apply parallelizing tools (Michelle knows how).

Goals of this particular model:
To explore making this multilocus and replicated, we need a bigger test dataset. 
bundles of 12 replicate single locus runs are similar to our real 12-locus multi-locus dataset. We also want multiples of those to build up a distribution of possible outcomes. This means something like 20 reps * 12 loci = 240 MetaPopGen single locus replicates.  Another thing that would be better in a play dataset would be if there was some differentiation. Flat structure makes it hard to follow plots sometimes. Remove all migration and see if that enables some differentiation. Also run a little longer (might be needed to build up differentiation). (All comparative statements based on work on testV8 so far).

3/16/17 and 3/22/17 V2 attempts failed to save final results (crashing computer/failing).

Saving parameters, results and log to these files:
running_V3_4_5_17_240rep_no_migration_parallel_parameters.Rdata
running_V3_4_5_17_240rep_no_migration_parallel_results.Rdata
running_V3_4_5_17_240rep_no_migration_parallel_log.txt

Working here because network planned overnight would affect saving processes
C:/Users/labguest/Documents/Temporary_MetaPopGen_work

Later move to usual directory:
Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen

Overview of this model 
3 populations
50,000 years
240 replicates
5000 kappa juvenile carrying capacity
0.3 juvenile survival
0.9 flowering survival
2 age classes, 1 and 2 years, rosettes and flowering plants. 
0 overall migration rate (0 exchanged between each pair)
1e-5 mutation rate

full details in script here below as well as saved parameter and data files.

*************************************************************************************

MetaPopGen publication
https://www.ncbi.nlm.nih.gov/pubmed/25585533

Access:
https://sites.google.com/site/marcoandrello/metapopgen

More instructions:
?MetaPopGen

*************************************************************************************
Install MetaPopGen
If not already installed on working machine, download from here:
https://sites.google.com/site/marcoandrello/metapopgen

And point path below to the downloaded software.
```{r}
install.packages("BiasedUrn")
install.packages("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/MetaPopGen.0.0.4.tar.gz",
                 type = "source", repos=NULL)
```

*************************************************************************************
If already installed, start here.
Set up: clear history and make sure you're in a good working directory (files will be generated), load program.
```{r}
rm(list=ls())
getwd()
library(MetaPopGen)
```

*************************************************************************************
Populate parameter values and initial populations

# Part 1: Creating the initial dataset
```{r}
n <- 3		# Number of demes - 3 populations
z <- 2		# Number of ages, first year rosettes, second year flowering plants.
l <- 10		# Number of alleles -  observations average 8.5/locus; model 10
m <- l*(l+1)/2		# Number of genotypes = diploid combinations of alleles l(l+1)/2
tmax <- 50000		# Number of years of simulation

	# Initialize matrices that will hold survival and reproduction rates
  # using the correct sizes for each dimension
  # Dimensions = Genotypes x Demes x Ages x Time
		sigma <- array(NA,dim=c(m,n,z,tmax)) #survival 
		phi_M <- array(0,dim=c(m,n,z,tmax)) #male fecundity
		phi_F <- array(0,dim=c(m,n,z,tmax)) #female fecundity
  
		# Important to track order of events for interpreting this: 
	# 1) survival - multiply input  by survival, removes some rosettes and flowering plants.
	# 2) reproduction - flowering plants produce seed
	# 3) recruitment - some seed becomes rosettes, rosettes become flowers, flowers  removed
	# N1 is before you start any thing, N2 is after one cycle of this.
		
		# This assigns values to the variables
		sigma[,,1,] <- 0.3 #survival is 30% for rosettes 
		sigma[,,2,] <- 0.9 #survival is 90% for flowering
		phi_M[,,2,] <- 10000 #flowering males make 10,0000 pollen grains each
		phi_F[,,2,] <- 1000 #flowering females make 500 ovules each ~seed count

	# Seed Dispersal matrix
	# seeds move around before sprouting into rosettes in recruitment phase
	m.m <- 0 # migration no one goes anywhere
	perpop <- m.m/(n-1)
	delta <- matrix(rep(perpop,n^2),nrow=n,ncol=n)
	diag(delta)<-rep((1-m.m),n)
	
	# Initial population structure:
	# flat - all genotypes, all demes have the same number of individuals; different number of rosettes and flowering plants.
	# Initialize array for initial number of individuals for [genotypes * demes * ages]
		N1 <- array(NA,dim=c(m,n,z))		
		# different starting n individuals for different age classes
		# entering values that are per population/deme, per genotype
		nrosettes = 100 #lots of rosettes per deme per genotype
		nflowers = 10 #some flowering
    N1[,,1] <- nrosettes #fill in value across all genotypes, across all demes, age class 1
    N1[,,2] <- nflowers #fill in value across all genotypes, across all demes, age class 2
    
    
	# Mutation matrix - uses max alleles at locus to model k-alleles mutation
    overall_mu <- 1e-5 #likelihood of an allele mutating to any another allele
    not_mutating <- 1-overall_mu
    per_allele_mu <- overall_mu/(l-1) #there are l-1 nonself alleles to mutate to, each with equal likelihood 
    #** could change this to model stepwise - 50% chance of moving up or down 1,with careful handling of edges.
		mu <- array(per_allele_mu,dim=c(l,l)) # fill in array with per_allele_mu
		#then make the diagonals the not_mutating rate
		for (from_allele in 1:l){
		  for (to_allele in 1:l){
		    if (from_allele==to_allele) {
		      mu[from_allele, to_allele] <- not_mutating
		    }
		  }
		}

	# Maximum recruitment per deme: Juvenile Carrying Capacity
  # After reproducing, how many will make it into next year's starting Age 1
  # Flat kappas
	kappa0 <- array(5000,c(n,tmax)) #at most 5000 seedlings allowed per population.
  # excess of seed means that kappa usually reached, randomness near asymptote, not limited by seed
		
#Sample in time
#How many years between taking samples for later analysis? (too big to save all years)
interval <- 500 # sample every n years

#build a list of sampling generations based on this interval.
samplemid <- seq.int(interval, tmax-interval, interval)
samplegens <- c(1, samplemid, tmax)
print("Sampling times")
print(samplegens)
nt <- length(samplegens)
print(sprintf("Collecting %s time points", nt))
# Save parameter set
filepath <- "C:/Users/labguest/Documents/Temporary_MetaPopGen_work/running_V3_4_5_17_240rep_no_migration_parallel_parameters.Rdata" 

save(nt,samplegens,n,z,l,m,tmax,sigma,N1,phi_M,phi_F,mu,delta,kappa0,tmax, file=filepath)

```
# Examine set up created above
See how N1 came out - is the number of individuals set up to be different across age classes? Glance at mutation and migration matrices to make sure they are sensible.
```{r}
print("N1 Dimensions = genotype * deme * age class")
print(dim(N1)) # genotypes x  demes x  age classes
print("Genotype 1, deme 1, age 1 and 2")
print(N1[1,1,]) #look at 1 genotype, 1 deme, all age classes
print("Genotype 50, deme 3, age 1 and 2")
print(N1[50,3,]) #look at 1 genotype, 1 deme, all age classes
print(sprintf("%s rosettes per deme, %s flowering plants per deme to start", sum(N1[,1,1]), sum(N1[,1,2])))
print("print mutation rate matrix, between alleles")
print(mu)
print("print migration rate matrix, between demes")
print(delta)
```


*************************************************************************************

#####################################################################################
# Perform the simulation
#####################################################################################

	
```{r}
# Option: load saved workspace (starting from here instead of set up above)
#rm(list=ls())
#filepath <- "C:/Users/labguest/Documents/Temporary_MetaPopGen_work/running_V3_4_5_17_240rep_no_migration_parallel_parameters.Rdata" 
#library(MetaPopGen)
#load(filepath)

#gather tools for parallelizing
library(doParallel)
library(parallel)
library(abind)
library(foreach)


# Working from variables set up above (instead of reloading saved parameter file)

# Simulation: 20 reps each for 12 loci
nrepl <- 240

#Make a log file  
logfile <- "C:/Users/labguest/Documents/Temporary_MetaPopGen_work/running_V3_4_5_17_240rep_no_migration_parallel_log.txt"
writeLines(c(""), logfile )

#Set up parallel abilities
cl <- makeCluster(3)
registerDoParallel(cl)

	# Run the simulations
t0 = Sys.time()#note the time at the beginning
Ns_list <-	foreach(replicateindex = 1:nrepl,
              .packages = "MetaPopGen") %dopar% {
    t0i <- Sys.time() #beginning for this replicate
    N<-sim.metapopgen.monoecious(
    input.type="array",
    N1 = N1, sigma = sigma,
    phi_M = phi_M, phi_F = phi_F,
    mu = mu, delta = delta, kappa0 = kappa0,
    T_max = tmax,
    save.res = F)

		#store a time sample of N, after calculating a full replicate, to an array 
  	Nt <- array(NA, dim=c(m,n,z,nt))
		for (timepoint in 1:nt) {
  	  print(sprintf("Grabbing time point number %s of %s", timepoint, nt))
  	  whichyear <- samplegens[timepoint]
  		print(sprintf("Timepoint %s corresponds to year %s", timepoint, whichyear))
  		print(dim(N))
  		Nt[,,,timepoint] <- N[,,,whichyear]
		}
  	tfi <- Sys.time()#end for this replicate
  	
  	#report progress, into log file.
    saverline <- sprintf("Done with replicate %s of %s, which took %s minutes. %s hours since start.", replicateindex, nrepl,
                        format(as.numeric(tfi-t0i,units="mins"), digits=3),
                        format(as.numeric(tfi-t0,units="hours"), digits=3))
    write(saverline,file=logfile,append=TRUE)
    Nt
	}

# important step to clean up at end of parallelizing and go back to normal		
stopCluster(cl)

# foreach makes a list of datasets, one for each replicate
# convert to 5d array (replicates, genotype, deme, ageclass, time)
Ns <- abind(Ns_list, along=0)

#Save!!!
results_filepath <- "C:/Users/labguest/Documents/Temporary_MetaPopGen_work/running_V3_4_5_17_240rep_no_migration_parallel_results.Rdata" 
save(Ns,sigma,N1,phi_M,phi_F,mu,delta,kappa0,tmax,nrepl,file=results_filepath)
```		

