---
title: "Building MetaPopGen for Phacelia"
author: "Jessie Berta-Thompson"
date: "February 15, 2017"
output: html_document
---

Purpose of this version:
Starting from  simple Island model example, setting up options that get closer to phacelia (with annotation), and running very simple ones to understand how steps work and output.

Set up: clear history and make sure you're in a good working directory (files will be generated), load program.
```{r}
rm(list=ls())
getwd()
library(MetaPopGen)
#setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen")
```


# Part 1: Creating the initial dataset

```{r}
nind <- 20		# Initial number of individuals per age-class and genotype
n <- 2		# Number of demes
z <- 3		# Number of age classes
l <- 3		# Number of alleles
m <- 6		# Number of genotypes
tmax <- 5		# Number of years of simulation

	# Initialize matrices that will hold survival and reproduction rates
	# This initialises the variables using the correct sizes for each dimension
  # Dimensions = Genotypes x Demes x Ages x Time
		sigma <- array(NA,dim=c(m,n,z,tmax)) #survival 
		phi_M <- array(0,dim=c(m,n,z,tmax)) #male fecundity
		phi_F <- array(0,dim=c(m,n,z,tmax)) #female fecundity

	# This assigns values to the variables
		sigma[,,,] <- 1 #survival is 100%
		phi_M[,,,] <- 1 #Males make 1 pollen grain each
		phi_F[,,,] <- 2 #Females make 2 ovules

	# Dispersal matrix
	m.m <- 0.1
	delta <- matrix(rep(m.m/(n-1),n^2),nrow=n,ncol=n)
	diag(delta)<-rep((1-m.m),n)
#seeds (next year's age class 1 move around before settled)
	
	# Initial population structure
	# Genotypes number are assigned at random, by drawing the number of individuals of each genotype from a uniform distribution between [0,nind/m]
		N1 <- array(NA,dim=c(m,n,z))		# Initial number of individuals for [genotypes * demes * ages]

		for (deme in 1 : n){
			for (age in 1 : z) {
				for (genotype in 1 : m) {
					N1[genotype,deme,age] <- nind 
					#seems like it's just assigning the number to each - not drawn from, but assigned.
				}
			}
		}

	# Mutation matrix
		mu <- array(1e-6,dim=c(l,l))
		mu[1,1] <- 1 - mu[2,1]
		mu[2,2] <- 1 - mu[1,2]
  #2 alleles in diploid genotype - each can stay the same or change. 
		
		
	# Maximum recruitment
		kappa0 <- array(Inf,c(n,tmax))
  #Juvenile Carrying Capacity = infinity for now; after reproducting, how many age class 1 will make it to next year's starting Age 1
		
		
	# Save dataset
		save(sigma,N1,phi_M,phi_F,mu,delta,kappa0,tmax,
		     file="Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/test_V1_2_15_17.RData")
```

#############################################################################################
# Perform the simulation
#############################################################################################

	library(MetaPopGen)
```{r}
	# Clear the workspace
		rm(list=ls())

	# Load dataset (created in Part 1)
		load("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/test_V1_2_15_17.RData")

	# Simulation
		nrepl <- 1


	# Create the array for fst
		fst <- array(NA,dim=c(nrepl,tmax))
	
	# Run the simulations
	for (i in 1 : nrepl) {
	  
	  cat("Replicate",i,"\n") # Just to print on screen the advancement of the simulation
	  
	  N<-sim.metapopgen.monoecious(
		input.type="array",
		N1 = N1, sigma = sigma,
		phi_M = phi_M, phi_F = phi_F,
		mu = mu, delta = delta, kappa0 = kappa0,
		T_max = tmax,
		save.res = F)
		
		flush.console()
		
		# Calculate Fst
		print("Calculating fst...")
		for (t in 1 : tmax){
			fst[i,t] <- fst.global.monoecious(N,t)
		}
		

	}

```		

Working to understand output.
```{r}
#N is a new object 6 * 2 * 3 * 5 [genotypes * demes * ages * times]
print(dim(N))

#look at just one deme, each time individually - should be a matrix with genotypes vs age classes
year1 = N[,2,,1] #everything for time 1
year2 = N[,2,,2] #everything for time 2
year3 = N[,2,,3] #everything for time 2
year4 = N[,2,,4] #everything for time 2
year5 = N[,2,,5] #everything for time 2

print('Year 1')
print(year1)

print('Year 2')
print(year2)

print('Year 3')
print(year3)

print('Year 4')
print(year4)

print('Year 5')
print(year5)



```

Interpret these results:
year 1 to year 2:
only age class 1 changed.
some went up, some went down.
New ones created(with randomness) and no old ones stay in age class 1
maybe old age class 1 moved up to 2, old age class 2 moved up to 3
moving up might happen in first step or might in last step? which value gets multiplied to govern reproduction?




Calculations
```{r}
# Meand and standard deviation of FST
mean(fst[,100:tmax]) # because reached stable Fst after ca 50 generations?
sd(fst[,100:tmax])

# Expected value for FST
Ne <- 3000
m.m <- 0.1
fst.exp <- 1/(1 + 4 * Ne * (m.m))
```
