---
title: "Running MetaPopGen v2"
author: "Jessie Berta-Thompson"
date: "March 16, 2017; March 22, 2017"
output: html_document
---

Purpose of this version, immediate tasks:
To explore making this multilocus and replicated, we need a bigger test dataset. 
bundles of 8 replicate single locus runs are similar to our real 8-locus multi-locus dataset. We also want multiples of those to build up a distribution of possible outcomes. This means something like 20 reps * 8 loci = 160 MetaPopGen single locus replicates. 

Another thing that would be better in a play dataset would be if there was some differentiation. Flat structure makes it hard to follow plots sometimes. Remove all migration and see if that enables some differentiation. Also run a little longer (might be needed to build up differentiation). (All comparative statements based on work on testV8 so far).

MetaPopGen publication
https://www.ncbi.nlm.nih.gov/pubmed/25585533

3/16/17 version failed to save final results (not sure why). Try again on more powerful computer, giving the parameter choices a once over and checking saving file names.

Saving parameters and results to these files:
running_V2_3_22_17_160_rep_no_migration_parameters.Rdata
running_V2_3_22_17_160_rep_no_migration_results.Rdata

overview of this practice dataset (running V2)
3 populations
50,000 years
160 replicates
5000 kappa juvenile carrying capacity
0.3 juvenile survival
0.9 flowering survival
2 age classes, 1 and 2 years, rosettes and flowering plants. 
0 overall migration rate (0 exchanged between each pair)
1e-5 mutation rate

full details in script below as well as saved parameter and data files.

*************************************************************************************

List here thoughts for future work as they come up:

 - Eventually will continue exploring demographic stochasticity of bad years (make a version that's stochastic across replicates), but for now behavior of that one was odd - just work with a flatter demography while building up tools. Too big of swings in last attempt at bad years. 

 - Population age structure set by Kappa and survival rates. build a mathematical tool to pick those values based on more intuitive input parameters in order to create realistic population structure (current one a little flower-heavy)
 
- write generic code for making print tables (e.g. tabs, padding ). some of this in Rmetasim code already. for visualizing parameter choices nicely/fast comprehension.

- Add parallel tools to get faster run times

- Thing of replicates as replicates * loci for modelling multilocus (add parameter)

- Add a time report to printed statement as each replicate is complete

*************************************************************************************
Install MetaPopGen
If not already installed on working machine, download from here:
https://sites.google.com/site/marcoandrello/metapopgen

And point path below to the downloaded software.
```{r}
install.packages("BiasedUrn")
install.packages("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/MetaPopGen.0.0.4.tar.gz",
                 type = "source", repos=NULL)
```

*************************************************************************************
If already installed, start here:
Set up: clear history and make sure you're in a good working directory (files will be generated), load program.
```{r}
rm(list=ls())
getwd()
library(MetaPopGen)
```

*************************************************************************************
Populate parameter values and initial populations

# Part 1: Creating the initial dataset
```{r}
n <- 3		# Number of demes - 3 populations
z <- 2		# Number of ages, first year rosettes, second year flowering plants.
l <- 10		# Number of alleles -  observations average 8.5/locus; model 10
m <- l*(l+1)/2		# Number of genotypes = diploid combinations of alleles l(l+1)/2
tmax <- 50000		# Number of years of simulation

	# Initialize matrices that will hold survival and reproduction rates
  # using the correct sizes for each dimension
  # Dimensions = Genotypes x Demes x Ages x Time
		sigma <- array(NA,dim=c(m,n,z,tmax)) #survival 
		phi_M <- array(0,dim=c(m,n,z,tmax)) #male fecundity
		phi_F <- array(0,dim=c(m,n,z,tmax)) #female fecundity
  
		# Important to track order of events for interpreting this: 
	# 1) survival - multiply input  by survival, removes some rosettes and flowering plants.
	# 2) reproduction - flowering plants produce seed
	# 3) recruitment - some seed becomes rosettes, rosettes become flowers, flowers  removed
	# N1 is before you start any thing, N2 is after one cycle of this.
		
		# This assigns values to the variables
		sigma[,,1,] <- 0.3 #survival is 30% for rosettes 
		sigma[,,2,] <- 0.9 #survival is 90% for flowering
		phi_M[,,2,] <- 10000 #flowering males make 10,0000 pollen grains each
		phi_F[,,2,] <- 1000 #flowering females make 500 ovules each ~seed count

	# Seed Dispersal matrix
	# seeds move around before sprouting into rosettes in recruitment phase
	m.m <- 0 # migration no one goes anywhere
	perpop <- m.m/(n-1)
	delta <- matrix(rep(perpop,n^2),nrow=n,ncol=n)
	diag(delta)<-rep((1-m.m),n)
	
	# Initial population structure:
	# flat - all genotypes, all demes have the same number of individuals; different number of rosettes and flowering plants.
	# Initialize array for initial number of individuals for [genotypes * demes * ages]
		N1 <- array(NA,dim=c(m,n,z))		
		# different starting n individuals for different age classes
		# entering values that are per population/deme, per genotype
		nrosettes = 100 #lots of rosettes per deme per genotype
		nflowers = 10 #some flowering
    N1[,,1] <- nrosettes #fill in value across all genotypes, across all demes, age class 1
    N1[,,2] <- nflowers #fill in value across all genotypes, across all demes, age class 2
    
    
	# Mutation matrix - uses max alleles at locus to model k-alleles mutation
    overall_mu <- 1e-5 #likelihood of an allele mutating to any another allele
    not_mutating <- 1-overall_mu
    per_allele_mu <- overall_mu/(l-1) #there are l-1 nonself alleles to mutate to, each with equal likelihood 
    #** could change this to model stepwise - 50% chance of moving up or down 1,with careful handling of edges.
		mu <- array(per_allele_mu,dim=c(l,l)) # fill in array with per_allele_mu
		#then make the diagonals the not_mutating rate
		for (from_allele in 1:l){
		  for (to_allele in 1:l){
		    if (from_allele==to_allele) {
		      mu[from_allele, to_allele] <- not_mutating
		    }
		  }
		}

	# Maximum recruitment per deme: Juvenile Carrying Capacity
  # After reproducing, how many will make it into next year's starting Age 1
  # Flat kappas
	kappa0 <- array(5000,c(n,tmax)) #at most 5000 seedlings allowed per population.
  # excess of seed in system means that kappa usually reached, randomness near asymptote, not limited by seed
		
	# Add good and bad years:
	# this part needs some work, once evaluative tools up and running
	#	kappas <- c(500, 5000) 
	#number of seedlings in a bad year, number of seedlings in a good year
	#	probs <- c(0.2, 0.8) # probability of bad year, good year
	#	qualities <- vector(length=tmax)
	#	kappa0 <- array(NA, c(n, tmax))
	#	for (tp in 1:tmax) {
	#	  kappa0[,tp] <- sample(kappas, 1, prob=probs)
	#	}
		#this makes the same bad years across replicates - might be useful later to put randomness across reps.

	#Sample in time
#How many years between taking samples for analysis?
interval <- 200 # sample every n years

#build a list of sampling generations based on this interval.
samplemid <- seq.int(interval, tmax-interval, interval)
samplegens <- c(1, samplemid, tmax)
print("Sampling times")
print(samplegens)
nt <- length(samplegens)
print(sprintf("Collecting %s time points", nt))
# Save parameter set
filepath <- "Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/running_V2_3_22_17_160_rep_no_migration_parameters.Rdata" 

save(nt,samplegens,n,z,l,m,tmax,sigma,N1,phi_M,phi_F,mu,delta,kappa0,tmax, file=filepath)

```

See how N1 came out - is the number of individuals set up to be different across age classes?
```{r}
print("N1 Dimensions = genotype * deme * age class")
print(dim(N1)) # genotypes x  demes x  age classes
print("Genotype 1, deme 1, age 1 and 2")
print(N1[1,1,]) #look at 1 genotype, 1 deme, all age classes
print("Genotype 50, deme 3, age 1 and 2")
print(N1[50,3,]) #look at 1 genotype, 1 deme, all age classes
print(sprintf("%s rosettes per deme, %s flowering plants per deme to start", sum(N1[,1,1]), sum(N1[,1,2])))

```

#see mutation matrix came out
```{r}
print("print mutation rate matrix, between alleles")
print(mu)
print("print migration rate matrix, between demes")
print(delta)
```

*************************************************************************************

#####################################################################################
# Perform the simulation
#####################################################################################

	
```{r}
# Optional: load saved workspace (starting from here instead of set up above)
#rm(list=ls())
#filepath <- "Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/test_V5_ph_test_2_23_17.RData"
#library(MetaPopGen)
#load(filepath)

# Working from variables saved above.

	# Simulation
		nrepl <- 160

  # Create a place to store Ns as you go
		Ns <- array(NA, dim=c(nrepl, m, n, z, nt))
		
	# Create an array to store fst for each rep, each time point
		fst <- array(NA,dim=c(nrepl,tmax))
	
	# Run the simulations
	for (i in 1 : nrepl) {
	  
	  cat("Replicate",i,"\n") # Just to print on screen the advancement of the simulation
	  
	  N<-sim.metapopgen.monoecious(#check monoecious/dioecious
		input.type="array",
		N1 = N1, sigma = sigma,
		phi_M = phi_M, phi_F = phi_F,
		mu = mu, delta = delta, kappa0 = kappa0,
		T_max = tmax,
		save.res = F)
		
		flush.console() #ensures that display of output in the console is current, even if output buffering is on.
		
		#store a time sample of N, after calculating a replicate, to an array 
  	for (timepoint in 1:nt) {
  	  print(sprintf("Grabbing time point number %s of %s", timepoint, nt))
  	  whichyear <- samplegens[timepoint]
  		print(sprintf("Timepoint %s corresponds to year %s", timepoint, whichyear))
  		print(dim(N))
  		Ns[i,,,,timepoint] <- N[,,,whichyear]
  		print(dim(Ns))
  		}
		
		
		# Calculate global Fst for replicate across timepoints
		print("Calculating fst...")
		for (t in 1 : tmax){
			fst[i,t] <- fst.global.monoecious(N,t)
		}
	}

filepath <- "Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/MetaPopGen/running_V2_3_22_17_160_rep_no_migration_results.Rdata" 
save(Ns,fst,sigma,N1,phi_M,phi_F,mu,delta,kappa0,tmax,nrepl,file=filepath)
```		
