---
title: "Plotting_results_of_strataG_calculations_for_real_phacelia_dataset"
author: "Jessie Berta-Thompson"
date: "May 16, 2017"
output: html_document
---
*************************************************************************************
## Set up
Clear history and make sure you're in the right working directory,  load in strataG results.
```{r}
rm(list=ls())
getwd()
#then setwd() to where you want to be if wrong.
```

```{r}
#load in results for straight calculations using strataG on phacelia microsatellite data.
load("full_data_strataG_results_5_4_17.RData")
```

****************************************************************************************************

# Create a dataframe from pairwise by location Fst
with real data, melt it, and build up collection of plot-relevant factors, with meaningfully ordered and sorted levels.

```{r}
#Extract data and labels from strataG result object
fst <- as.numeric(real_by_loc_pT$result$Fst)
pair_label <- real_by_loc_pT$result$pair.label

#Bundle those into a dataframe
fst.df <- data.frame(Pair=pair_label, Fst = fst)

#Extract the two sites from the label as separate quantities and add to dataframe as additional factors
first_site<- unlist(lapply(pair_label, function(compstring) {
  halves <- strsplit(compstring, "v. ")
  messy_first <- halves[[1]][[1]]
  tidying <- strsplit(messy_first, " ")
  tidied <- tidying[[1]][[1]][[1]]
}))

second_site<- unlist(lapply(pair_label, function(compstring) {
  halves <- strsplit(compstring, "v. ")
  messy_second <- halves[[1]][[2]]
  tidying <- strsplit(messy_second, " ")
  tidied <- tidying[[1]][[1]][[1]]
}))

fst.df$Site1 <- factor(first_site)
fst.df$Site2 <- factor(second_site)


# Add Species designations for first and second of comparison pairs.
scullyi <- as.character(c("HN", "HS", "F"))
formosula <- as.character(c("CG", "D", "P","RN", "WS", "WA", "B", "V", "R", "AN", "AS"))
gina_glenneae <- as.character(c("T"))


species1 <- unlist(lapply(first_site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

species2 <- unlist(lapply(second_site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

fst.df$Species1 <- factor(species1)
fst.df$Species2 <- factor(species2)

#Add Counties (equivalent to species; but perhaps we'd want to label things geographically)
county1 <- unlist(lapply(species1, function(sp) {
  if (sp == "scullyi") {
    co <- "Larimer"
  }
  else if (sp == "formosula") {
    co <- "Jackson"
  }
  else if (sp == "gina-glenneae") {
    co <- "Grand"
  }
  else {
    print("uh-oh. mapping fail.")
  }
}))

county2 <- unlist(lapply(species2, function(sp) {
  if (sp == "scullyi") {
    co <- "Larimer"
  }
  else if (sp == "formosula") {
    co <- "Jackson"
  }
  else if (sp == "gina-glenneae") {
    co <- "Grand"
  }
  else {
    print("uh-oh. mapping fail.")
  }
}))

fst.df$County1 <- factor(county1)
fst.df$County2 <- factor(county2)


#Add combinations of species and counties
species_pair <- unlist(lapply(1:length(species1), function(i) {
  pair <- paste(species1[i], "vs.", species2[i]) 
}))

fst.df$SpeciesComparison <- factor(species_pair, levels = c("formosula vs. gina-glenneae", "formosula vs. formosula", "scullyi vs. formosula", "scullyi vs. scullyi", "scullyi vs. gina-glenneae"))

counties_pair <- unlist(lapply(1:length(species1), function(i) {
  pair <- paste(county1[i], "vs.", county2[i]) 
}))

fst.df$CountyComparison <- factor(counties_pair, levels = c("Jackson vs. Grand","Jackson vs. Jackson","Larimer vs. Jackson", "Larimer vs. Larimer", "Larimer vs. Grand"))

#Add one more, sorting by whether pair in a comparison is from the same county/species or different.
same_or_mixed <- unlist(lapply(1:length(species1), function(i) {
  if ( species1[i] == species2[i] ) {
    result <- "same species" #if you prefer county/species language, change these strings
  }
  else {
    result <- "different species"
  }
  return(result)
}))

fst.df$SameOrDiff <- factor(same_or_mixed)


#grab p-values
Fst_p_val <- real_by_loc_pT$result$Fst.p.val

fst.df$Fstpval <- Fst_p_val

print(fst.df)
```


## Color by species comparison, order original alphabetical order
```{r}
#So organized! Time to plot! In different orders! Other stats too, not just Fst - functionitize and repeat when done (enter stat) or add to data frame for combinations. 
library(ggplot2)
pdf(file = "real_data_pairwise_Fst_color_by_comparison_species_alpha_order_5_16_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=fst.df, mapping=aes(x = fst.df$Pair, y=fst.df$Fst,  color = fst.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites: alphabetical order")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))
```

## Reorder according to Fst value, keeping same color scheme by species comparison
```{r}
#change factor levels order to sort category-style x-axis. 

#make a copy  of the data frame to manipulate.
sorted_by_Fst.df <- fst.df 

#grab relevant pieces
unsorted_Fsts <- sorted_by_Fst.df$Fst
unsorted_Names <- sorted_by_Fst.df$Pair

#make the Fsts vector into a numeric with names
names(unsorted_Fsts) <- unsorted_Names

#sort Fst values
sortedFsts <- sort(unsorted_Fsts)

#see what that did to the names (good, out of order)
print(names(sortedFsts))
pairs_sorted_by_Fst_value <- names(sortedFsts)

#change order of levels
sorted_by_Fst.df$Pair <- factor(sorted_by_Fst.df$Pair, levels = pairs_sorted_by_Fst_value)

#file name for image:
pdf(file = "real_data_pairwise_Fst_color_by_comparison_species_Fst_order_5_16_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=sorted_by_Fst.df, mapping=aes(x = sorted_by_Fst.df$Pair, y=sorted_by_Fst.df$Fst,  color = sorted_by_Fst.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites: ordered by Fst")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))
```

## Reorder according to species comparison, keeping same color scheme by species comparison
```{r}
#make a copy  of the data frame sorted by species comparison.
sorted_by_species_comparison.df <- fst.df[order(fst.df$SpeciesComparison, fst.df$Fst),]

#see what it looks like:
print(sorted_by_species_comparison.df$Pair) #in an interesting order.

#change levels to match order in new sort.
sorted_by_species_comparison.df$Pair <- factor(sorted_by_species_comparison.df$Pair, levels = sorted_by_species_comparison.df$Pair)


#file name for image:
pdf(file = "real_data_pairwise_Fst_color_by_comparison_species_species_comparison_then_fst_order_5_16_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=sorted_by_species_comparison.df, mapping=aes(x = sorted_by_species_comparison.df$Pair, y=sorted_by_species_comparison.df$Fst,  color = sorted_by_species_comparison.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites: ordered by species comparison and Fst")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))
```



#others
add all parts of results to dataframe. 
alpha p value
plot just real values with color scheme, add element on p value (alpha)
sort by real value: get mean value with names, sort, change factor order. 
Plot Fsts from real data in same color scheme/styles.
Also: Sort by species comparison, then by value.
Also: Sort by value, color by within or between species.

Color by same or not:
fill = melted$samenot

Color by county compare language instead of species comparison

Could order or color by physical distance. 

5 categories: species comparisons with means, sd, ranges, real data plotted.
```{r}

ggplot(fst.df, aes(fst.df$SpeciesComparison, fst.df$Fst)) +
  geom_boxplot(outlier.colour = NA) + 
  geom_point(position = position_jitter(width = 0.2), mapping = aes(color = fst.df$SpeciesComparison)) +
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites\nsorted by species comparison\n")+
  xlab("")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("real_data_pairwise_Fst_by_comparison_species_species_boxplot_5_16_17.pdf", width = 9, height = 6, units = "in")

  
  







```


