---
title: "plotting_strataG_results_for_subsampling_phacelia_dataset_5_25_17.Rmd"
author: "Jessie Berta-Thompson"
date: "May 16, 2017"
output: html_document
---
*************************************************************************************
## Set up
Clear history and make sure you're in the right working directory,  load in subsampling strataG results.
```{r}
rm(list=ls())
here <- getwd()
print(here)
#then setwd() to where you want to be if wrong.
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Subsampling")
#look for files of interest
#list.files(here)
result_files <- list.files(here, pattern="_Fst_calcs_subsample_1000x_by_location_5_16_17.RData")
```

Start some manipulations to organize the files
```{r}
#extract n from file name
nvalues_from_file <- strtoi(unlist(lapply(result_files, function(filename) {
  strsplit(filename, "n_")[[1]][[1]]
})))

inputinforaw <- data.frame(files=result_files, n=nvalues_from_file, stringsAsFactors = FALSE)
sortorder <- order(inputinforaw["n"])
inputinfosorted <- inputinforaw[sortorder,]
row.names(inputinfosorted) <- c(1:dim(inputinfosorted["n"])[1])
```

figure out how to load and store the results
```{r}
simplevector <- as.character(inputinfosorted$files)
print(data.class(simplevector))
print(length(simplevector))


list_of_results <- lapply(simplevector, function(filename){
  print(filename) #progress monitor + ordercheck. 
  load(filename) #this loads a variable called thousand reps
  return(thousandreps)
})


```


Figure out how to access whatever that created
```{r}
#The whole thing is a list of 25 objects (the 25 ns done so far)
print(data.class(list_of_results))
print(length(list_of_results))

#Each of those items is a list of length 1
print(data.class(list_of_results[1]))
print(length(list_of_results[1]))

#Each of those contains a list of 1000 (the 1000 replicates of the n sampling)
print(data.class(list_of_results[[1]]))
print(length(list_of_results[[1]]))

#Each of those 1000 objects is a list of length 1.
print(data.class(list_of_results[[1]][1]))
print(length(list_of_results[[1]][1]))

#Each of those 1000 lists contains a list of 3 elements, the strataG results. 
print(data.class(list_of_results[[1]][[1]]))
print(length(list_of_results[[1]][[1]]))

#Those three elements are: $result, $pair.mat, $null.dist
print(data.class(list_of_results[[1]][[1]]))
print(length(list_of_results[[1]][[1]]))
print(list_of_results[[1]][[1]])

#what we want lives in $result.
print(list_of_results[[1]][[1]]$result$pair.label)
print(list_of_results[[1]][[1]]$result$Fst)
print(list_of_results[[1]][[1]]$result$Fst.p.val)

```

Build a dataframe with full dataset in easy-access form. 
```{r}
ns <- as.integer(inputinfosorted$n)

#loop over ns
data.df <- do.call(rbind, lapply(1:length(ns), function(i) {
#loop over reps
  allreps_for_one_n <- lapply(1:length(list_of_results[[i]]), function(r) {
    #grab just one result - one n, one rep.
    onerep <- list_of_results[[i]][[r]]
    len <- length(onerep$result$Fst)
    column1 = rep(ns[i], len)
    column2 = rep(r, len)
    column3 = onerep$result$pair.label
    column4 = onerep$result$Fst
    column5 = onerep$result$Fst.p.val

    one.df <- data.frame(n = column1, 
                         rep = column2, 
                         pairs = column3,
                         Fst = column4, 
                         pval = column5)
    return(one.df)
  })
  onen.df <- do.call(rbind, allreps_for_one_n)
}))

```


To this dataframe add more factor descriptors flushing out pair label information
```{r}
#Extract the two sites from the pair as separate quantities and add to dataframe as additional factors

first_site<- unlist(lapply(as.character(data.df$pairs), function(compstring) {
  first <- strsplit(compstring, " ")[[1]][[1]]
}))
second_site<- unlist(lapply(as.character(data.df$pairs), function(compstring) {
  second <- strsplit(compstring, " ")[[1]][[4]]
}))

#check them out
print(length(first_site))
print(length(second_site))
print(first_site[1:5])
print(second_site[1:5])
print(first_site[493:582])
print(second_site[24567:24572])

data.df$site1 <- first_site
data.df$site2 <- second_site

```


```{r}
# Add Species designations for first and second of comparison pairs.
scullyi <- as.character(c("HN", "HS", "F"))
formosula <- as.character(c("CG", "D", "P","RN", "WS", "WA", "B", "V", "R", "AN", "AS"))
gina_glenneae <- as.character(c("T"))

species1 <- unlist(lapply(first_site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

species2 <- unlist(lapply(second_site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

#check them out
print(length(species1))
print(length(species2))
print(species1[1:5])
print(species2[1:5])
print(species1[493:582])
print(species2[24567:24572])

data.df$species1 <- species1
data.df$species2 <- species2

```


```{r}
#Add Counties (equivalent to species; but perhaps we'd want to label things geographically)
county1 <- unlist(lapply(species1, function(sp) {
  if (sp == "scullyi") {
    co <- "Larimer"
  }
  else if (sp == "formosula") {
    co <- "Jackson"
  }
  else if (sp == "gina-glenneae") {
    co <- "Grand"
  }
  else {
    print("uh-oh. mapping fail.")
  }
}))

county2 <- unlist(lapply(species2, function(sp) {
  if (sp == "scullyi") {
    co <- "Larimer"
  }
  else if (sp == "formosula") {
    co <- "Jackson"
  }
  else if (sp == "gina-glenneae") {
    co <- "Grand"
  }
  else {
    print("uh-oh. mapping fail.")
  }
}))

data.df$county1 <- county1
data.df$county2 <- county2
```


```{r}
#Add combinations of species and counties
species_pair <- unlist(lapply(1:length(species1), function(i) {
  pair <- paste(species1[i], "vs.", species2[i]) 
}))

data.df$speciespair <- factor(species_pair, levels = c("formosula vs. gina-glenneae", "formosula vs. formosula", "scullyi vs. formosula", "scullyi vs. scullyi", "scullyi vs. gina-glenneae"))
```

```{r}
counties_pair <- unlist(lapply(1:length(species1), function(i) {
  pair <- paste(county1[i], "vs.", county2[i]) 
}))

data.df$countypair <- factor(counties_pair, levels = c("Jackson vs. Grand","Jackson vs. Jackson","Larimer vs. Jackson", "Larimer vs. Larimer", "Larimer vs. Grand"))
```

```{r}
#Add one more, sorting by whether pair in a comparison is from the same county/species or different.
same_or_mixed <- unlist(lapply(1:length(species1), function(i) {
  if ( species1[i] == species2[i] ) {
    result <- "same species" #if you prefer county/species language, change these strings
  }
  else {
    result <- "different species"
  }
  return(result)
}))

data.df$samespecies <- factor(same_or_mixed)
```


```{r}
#load in results forstrataG Fst calculations on phacelia microsatellite data.
load("full_data_strataG_results_5_4_17.RData")
```

##Average over replicates
```{r}
#loop over values of n
mean_rep.df <- do.call(rbind, lapply(1:length(ns), function(i) {
    onevaluen.df <- data.df[data.df$n == ns[i],]

    #loop over instances of pairs
    one_n <- lapply(unique(data.df$pairs), function(pair) {
        one_n_one_pair.df <- onevaluen.df[data.df$pairs == pair,]      
        fstmean = mean(one_n_one_pair.df$Fst)
        fststdev = sd(one_n_one_pair.df$Fst)
        #same for pvalue
        pvalmean = mean(one_n_one_pair.df$)
        pvalstdev = sd(one_n_one_pair.df$Fst)
        #then build up a dataframe. 
    })
  
  
  
  
  
}))



```



##Start plotting

```{r}
#So organized! Time to plot! In different orders! Other stats too, not just Fst - functionitize and repeat when done (enter stat) or add to data frame for combinations. 
library(ggplot2)
pdf(file = "samp_pairwise_Fst_color_by_comparison_species_alpha_order_5_16_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=fst.df, mapping=aes(x = fst.df$Pair, y=fst.df$Fst,  color = fst.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites: alphabetical order")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))
```











old plotting stuff from plain dataset that might come in handy. 
********************************************************************************************

## Color by species comparison, order original alphabetical order
```{r}
#So organized! Time to plot! In different orders! Other stats too, not just Fst - functionitize and repeat when done (enter stat) or add to data frame for combinations. 
library(ggplot2)
pdf(file = "real_data_pairwise_Fst_color_by_comparison_species_alpha_order_5_16_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=fst.df, mapping=aes(x = fst.df$Pair, y=fst.df$Fst,  color = fst.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites: alphabetical order")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))
```

## Reorder according to Fst value, keeping same color scheme by species comparison
```{r}
#change factor levels order to sort category-style x-axis. 

#make a copy  of the data frame to manipulate.
sorted_by_Fst.df <- fst.df 

#grab relevant pieces
unsorted_Fsts <- sorted_by_Fst.df$Fst
unsorted_Names <- sorted_by_Fst.df$Pair

#make the Fsts vector into a numeric with names
names(unsorted_Fsts) <- unsorted_Names

#sort Fst values
sortedFsts <- sort(unsorted_Fsts)

#see what that did to the names (good, out of order)
print(names(sortedFsts))
pairs_sorted_by_Fst_value <- names(sortedFsts)

#change order of levels
sorted_by_Fst.df$Pair <- factor(sorted_by_Fst.df$Pair, levels = pairs_sorted_by_Fst_value)

#file name for image:
pdf(file = "real_data_pairwise_Fst_color_by_comparison_species_Fst_order_5_16_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=sorted_by_Fst.df, mapping=aes(x = sorted_by_Fst.df$Pair, y=sorted_by_Fst.df$Fst,  color = sorted_by_Fst.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites: ordered by Fst")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))
```

## Reorder according to species comparison, keeping same color scheme by species comparison
```{r}
#make a copy  of the data frame sorted by species comparison.
sorted_by_species_comparison.df <- fst.df[order(fst.df$SpeciesComparison, fst.df$Fst),]

#see what it looks like:
print(sorted_by_species_comparison.df$Pair) #in an interesting order.

#change levels to match order in new sort.
sorted_by_species_comparison.df$Pair <- factor(sorted_by_species_comparison.df$Pair, levels = sorted_by_species_comparison.df$Pair)


#file name for image:
pdf(file = "real_data_pairwise_Fst_color_by_comparison_species_species_comparison_then_fst_order_5_16_17.pdf", width = 12, height = 5)

ggplot() + 
  geom_point(data=sorted_by_species_comparison.df, mapping=aes(x = sorted_by_species_comparison.df$Pair, y=sorted_by_species_comparison.df$Fst,  color = sorted_by_species_comparison.df$SpeciesComparison), size=3 )+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites: ordered by species comparison and Fst")+
  xlab("Pairwise comparisons")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="top")+
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))
```



#others
add all parts of results to dataframe. 
alpha p value
plot just real values with color scheme, add element on p value (alpha)
sort by real value: get mean value with names, sort, change factor order. 
Plot Fsts from real data in same color scheme/styles.
Also: Sort by species comparison, then by value.
Also: Sort by value, color by within or between species.

Color by same or not:
fill = melted$samenot

Color by county compare language instead of species comparison

Could order or color by physical distance. 

5 categories: species comparisons with means, sd, ranges, real data plotted.
```{r}

ggplot(fst.df, aes(fst.df$SpeciesComparison, fst.df$Fst)) +
  geom_boxplot(outlier.colour = NA) + 
  geom_point(position = position_jitter(width = 0.2), mapping = aes(color = fst.df$SpeciesComparison)) +
  scale_colour_manual(values=c("green4", "royalblue", "darkorchid","#DC143C", "#FF9000" ))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Fst values for all pairwise comparisons between 15 sites\nsorted by species comparison\n")+
  xlab("")+
  ylab("Fst")+
  theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

ggsave("real_data_pairwise_Fst_by_comparison_species_species_boxplot_5_16_17.pdf", width = 9, height = 6, units = "in")

```


