---
title: "Reformatting Phacelia microsatellite data for BayesAss"
author: "Jessie Berta-Thompson"
date: "June 8, 2017"
output: html_document
---

## Set up and load in data
The input is a simplified version of microsatellite data, originally from Phfo_trunkated data_11-24-14, sheet of ??? spreadsheet.

```{r}
#clean up
rm(list=ls())
#current directory?
getwd() #if needed, move with setwd()

#load in data
rawtable <- read.table("simplified_dataset_for_R.txt", header=TRUE, sep = "\t")

#what format?
head(rawtable)
data.class(rawtable)
dim(rawtable)

```
## To this dataframe, add population designations for different ways of separating the sites.

```{r}
# By Species
scullyi <- as.character(c("HN", "HS", "F"))
formosula <- as.character(c("CG", "D", "P","RN", "WS", "WA", "B", "V", "R", "AN", "AS"))
gina_glenneae <- as.character(c("T"))

species <- unlist(lapply(rawtable$Site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% formosula) {
    answer <- "formosula"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

rawtable$Species <- species
rawtable$Species <- as.factor(rawtable$Species)

# See if it worked by looking at column names and spot checking a few
colnames(rawtable)
head(rawtable)
tail(rawtable)
```


```{r}
# Combine close pairs: pairs that break BayesAss program at 2/3 max migrant prior are combined
fine_alone <- as.character(c("F","CG","D","P","WA","R","T"))
hnhs <- as.character(c("HN", "HS"))
rnws <- as.character(c("RN", "WS"))
bv <- as.character(c("B", "V"))
anas <- as.character(c("AN", "AS"))

combine_close_pairs <- unlist(lapply(rawtable$Site, function(site) {
  if (site %in% fine_alone) {
    answer <- as.character(site)
  }
  else if (site %in% hnhs) {
    answer <- "HNHS"
  }
  else if (site %in% rnws) {
    answer <- "RNWS"
  }
  else if (site %in% bv) {
    answer <- "BV"
  }
  else if (site %in% anas) {
    answer <- "ANAS"
  }
  return(answer)
}))

rawtable$CombineClosePairs <- combine_close_pairs
rawtable$CombineClosePairs <- as.factor(rawtable$CombineClosePairs)

# See if it worked by looking at column names and spot checking
head(rawtable)
print(combine_close_pairs)
```

```{r}
# 2 formosula clusters
#by species, except splitting formosula in two, based on geography, clustering and Fst.

scullyi <- as.character(c("HN", "HS", "F"))
north_formosula <- as.character(c("CG", "D", "P","RN", "WS", "WA", "B", "V"))
south_formosula <- as.character(c("R", "AN", "AS"))
gina_glenneae <- as.character(c("T"))

two_formosula_clusters <- unlist(lapply(rawtable$Site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% north_formosula) {
    answer <- "NF"
  }
  else if (site %in% south_formosula) {
    answer <- "SF"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

rawtable$TwoFormoClust <- two_formosula_clusters
rawtable$TwoFormoClust <- as.factor(rawtable$TwoFormoClust)

# See if it worked by looking at column names and spot checking a few
rawtable[which(rawtable$Site == "AN"), ] #grabs rows for AN site
rawtable[which(rawtable$Site == "D"), ]
```

```{r}
# 3 formosula clusters
#by species, except splitting formosula in two, based on geography, clustering and Fst, and moving most different site, WA, to its own category.

scullyi <- as.character(c("HN","HS","F"))
west_formosula <- as.character(c("CG","D","P","RN","WS","B","V"))
alone <- as.character(c("WA"))
south_formosula <- as.character(c("R","AN","AS"))
gina_glenneae <- as.character(c("T"))

three_formosula_clusters <- unlist(lapply(rawtable$Site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% west_formosula) {
    answer <- "WF"
  }
  else if (site %in% alone) {
    answer <- as.character(site)
    print(answer)
  }
  else if (site %in% south_formosula) {
    answer <- "SF"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

rawtable$ThreeFormoClust <- three_formosula_clusters
rawtable$ThreeFormoClust <- as.factor(rawtable$ThreeFormoClust)

# See if it worked by looking at column names and spot checking a few
rawtable[which(rawtable$Site == "WA"), ] #grabs rows for AN site
rawtable[which(rawtable$Site == "D"), ]
```


```{r}
# 4 formosula clusters
#by species, except splitting formosula in two, based on geography, clustering and Fst, and moving most different sites, WA and CG, each to their own category.

scullyi <- as.character(c("HN","HS","F"))
central_formosula <- as.character(c("D","P","RN","WS","B","V"))
alone <- as.character(c("WA", "CG"))
south_formosula <- as.character(c("R","AN","AS"))
gina_glenneae <- as.character(c("T"))

four_formosula_clusters <- unlist(lapply(rawtable$Site, function(site) {
  if (site %in% scullyi) {
    answer <- "scullyi"
  }
  else if (site %in% central_formosula) {
    answer <- "CF"
  }
  else if (site %in% alone) {
    answer <- as.character(site)
  }
  else if (site %in% south_formosula) {
    answer <- "SF"
  }
  else if (site %in% gina_glenneae) {
    answer <- "gina-glenneae"
  }
  return(answer)
}))

rawtable$FourFormoClust <- four_formosula_clusters
rawtable$FourFormoClust <- as.factor(rawtable$FourFormoClust)

# See if it worked by looking at column names and spot checking a few
rawtable[which(rawtable$Site == "CG"), ] #grabs rows for AN site
rawtable[which(rawtable$Site == "D"), ]
```


## Rearrange to get to BayesAss input format
5 columns, separated by spaces:
indivID	popID	locID	allele1	allele2

In current form, loci are columns; rearrange so each locus in each individual is one row.
#Do some legwork to sort out new table properties and get locus names in useful form.
```{r}
## How long will the final dataframe be? (ind * loci)
# What are dimensions of current dataframe
dims <- dim(rawtable) 
data.class(dims)
length(dims)
print(dims)

#number of loci - should be 12 for this study. 
colnames(rawtable)
length(colnames(rawtable))
number_of_non_locus_columns = 7
nl <- (dims[2] - number_of_non_locus_columns)/2
print(nl)

#total number of individuals
ni <- dims[1]

#number of rows in dream dataframe = ind * loci
nr <- ni * nl
print(nr)

#extract locus names
ints <- sapply(rawtable, is.integer) #only integer data in dataframe is locus data. get only integers
locus_cols <- colnames(rawtable[,ints]) #get headers for only integers slice of dataframe.

#loci column names are in this format: "X2638...1" "X2638...2" 
clean_up_locus_names <- unlist(lapply(locus_cols, function(inputstring) {
  print(inputstring)
  #split at "." and take first piece.
  partway <- strsplit(inputstring, "\\.")[[1]][[1]]
  print(partway)
  #split at "X" and take second piece.
  finish <- strsplit(partway, "X")[[1]][[2]]
  return(finish)
})) 
print(clean_up_locus_names)

#grab every other one - now there are repeats. preserve order.
twox <- length(clean_up_locus_names)
onex <- twox/2
locus_names <- unlist(lapply(1:onex, function(i) {
  clean_up_locus_names[2*i]#always take second one.
}))
print(locus_names)
```


#For each desired output file in the 6 different population organizations, build a dataframe.

```{r}
CreateFile <- function(popsys) {
  #loop over rows of original dataframe (all individuals), and create rows of new one.
  converted.df <- do.call("rbind", lapply(c(1:ni), function(ind) { # loop over individuals ni (as index)
  input_row <- rawtable[ind,]
  one_ind_all_loci_list <- lapply (c(1:nl), function(loc){ 
    # loop over loci, producing a list nl long, with a row for each locus for that individual.
    onerow <- data.frame(indivID = input_row$Ind, 
                         popID = input_row[,popsys], 
                         locID =locus_names[loc], 
                         allele1 = input_row[1,2+(2*loc)-1], 
                         allele2 = input_row[1,2+(2*loc)])
    return(onerow)
  })
  #convert list of df rows to a df
  one_ind_all_loci.df <- do.call("rbind", one_ind_all_loci_list)
  return(one_ind_all_loci.df)
}))

filename <- paste("phacelia_bayesass_input_",popsys, ".txt", sep = "")
print(filename)
write.table(converted.df, filename, sep=" ", row.names=FALSE, col.names=FALSE, quote=FALSE)
}
```
Try using function to create files.
```{r}
CreateFile("Site")
CreateFile("Species")
CreateFile("CombineClosePairs")
CreateFile("TwoFormoClust")
CreateFile("ThreeFormoClust")
CreateFile("FourFormoClust")
```

## Finally, do all the same stuff but for a dataset without troublesome creek.

```{r}
NoTroubleTable <- rawtable[which(rawtable$Site != "T"),]

## How long will the final dataframe be? (ind * loci)
# What are dimensions of current dataframe
dims <- dim(NoTroubleTable) 

#number of loci - should be 12 for this study. 
number_of_non_locus_columns = 7
nl <- (dims[2] - number_of_non_locus_columns)/2
print(nl)

#total number of individuals
ni <- dims[1]
print(ni)

#number of rows in dream dataframe = ind * loci
nr <- ni * nl
print(nr)

#extract locus names
ints <- sapply(NoTroubleTable, is.integer) #only integer data in dataframe is locus data. get only integers
locus_cols <- colnames(NoTroubleTable[,ints]) #get headers for only integers slice of dataframe.

#loci column names are in this format: "X2638...1" "X2638...2" 
clean_up_locus_names <- unlist(lapply(locus_cols, function(inputstring) {
  #print(inputstring)
  #split at "." and take first piece.
  partway <- strsplit(inputstring, "\\.")[[1]][[1]]
  #print(partway)
  #split at "X" and take second piece.
  finish <- strsplit(partway, "X")[[1]][[2]]
  return(finish)
})) 
print(clean_up_locus_names)

#grab every other one - now there are repeats. preserve order.
twox <- length(clean_up_locus_names)
onex <- twox/2
locus_names <- unlist(lapply(1:onex, function(i) {
  clean_up_locus_names[2*i]#always take second one.
}))
print(locus_names)
```


#For each desired output file in the 6 different population organizations, build a dataframe.
```{r}
CreateFileNoTrouble <- function(popsys) {
  #loop over rows of original dataframe (all individuals), and create rows of new one.
  converted.df <- do.call("rbind", lapply(c(1:ni), function(ind) { # loop over individuals ni (as index)
  input_row <- NoTroubleTable[ind,]
  one_ind_all_loci_list <- lapply (c(1:nl), function(loc){ 
    # loop over loci, producing a list nl long, with a row for each locus for that individual.
    onerow <- data.frame(indivID = input_row$Ind, 
                         popID = input_row[,popsys], 
                         locID =locus_names[loc], 
                         allele1 = input_row[1,2+(2*loc)-1], 
                         allele2 = input_row[1,2+(2*loc)])
    return(onerow)
  })
  #convert list of df rows to a df
  one_ind_all_loci.df <- do.call("rbind", one_ind_all_loci_list)
  return(one_ind_all_loci.df)
}))

filename <- paste("phacelia_bayesass_input_",popsys, "_excludingTroublesomeCreek.txt", sep = "")
print(filename)
write.table(converted.df, filename, sep=" ", row.names=FALSE, col.names=FALSE, quote=FALSE)
}
```
Try using function to create files.
```{r}
CreateFileNoTrouble("Site")
CreateFileNoTrouble("Species")
CreateFileNoTrouble("CombineClosePairs")
CreateFileNoTrouble("TwoFormoClust")
CreateFileNoTrouble("ThreeFormoClust")
CreateFileNoTrouble("FourFormoClust")
```

```
