---
title: "Rmetasim_processing_saved_data_V4"
author: "Jessie Berta-Thompson"
date: "February 8, 2017"
output:
  html_document: default
  pdf_document: default
---
## immediate next steps for tomorrow. test piecemeal method for calculating and plotting, reorganize script so clear.
## add steps to save and load saved genind files (costly to rerun)
## implement sampling versions. 



## Version: using V6 2000 plants + seeds (2% plants total) + 5000 generations + 10 replicates, improve this script to plot more cleanly and explore more useful statistics. Main goal, finishing cleaning up counts plotting and then incorporate population genetics scripts, using same plotting framework.

## Structure of Rmetasim output files that this script parses
Each output file contains a replist, from one Rmetasim run, the structure of which is:
[[[rep1 generations sampled list][rep1 rland$individuals list of matrices]] [[rep2 generations sampled list][rep2 rland$individuals list of matrices]]]
replist is a list containing r objects, where r is the number of replicates, with a dataset for each replicate.
Each of the replicate objects is a list containing 2 objects, a list containing the year for generations sampled and a list containing rland$individuals matrices with genetic data for each generation sampled.
Sampling from nested lists:
This is the first replicate's (1) individual list (2) containing rland$individuals for all time points
replist[[1]][[2]]
This is the first replicate's (1) list of generation times (1) that correspond to samples in individual list
replist[[1]][[1]]

## Processing rmetasim output
Output datafiles from our Rmetasim runs contain a 'replist' list of lists with generations sampled and rland$individuals for each replicate.
```{r}
#clear environment
rm(list =ls())
#load desired dataset
load("/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen.Rdata") 

#usual path example
#load("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Rmetasim/replist_1_19_17_V6_seed_test_5000gen.Rdata")
```


## Extract some useful values about the dataset from its dimensions
```{r}
#How many replicates does this dataset contain?
nreps = length(replist)
print(sprintf("This dataset contains %s replicate simulations.", nreps))

#How many time samples were collected for each replicate? How many time samples maximum? (a given replicate can have less than the input length of run when extinctions occur)
nsamples_list <- list()
for (i in 1:length(replist)){
  nsamples <-length(replist[[i]][[1]])
  nsamples_list[i] <- nsamples}

maxsamples <- max(unlist(nsamples_list))
print(sprintf("At most, %s time samples were collected.", maxsamples))

#How many years of simulation does this represent?
nyears_list <- list()
for (i in 1:length(replist)){
  nyears <-replist[[i]][[1]][length(replist[[i]][[1]])]
  nyears_list[i] <- nyears}

maxyears <- max(unlist(nyears_list))
print(sprintf("At most, this simulation was run for %s years.", maxyears))

#what years were sampled (for the longest)?
index_of_longest <- which.max(unlist(nyears_list))
sample_years <- replist[[index_of_longest]][[1]]
print("These years were sampled:")
print(unlist(sample_years))

```


## Working on a more sophisticated & generalizable plotting structure
ggplot recommended for complicated plotting
takes a dataframe as input
```{r}
library("reshape2")
jpeg("/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_8_multicolor.jpg", width=600, height=400, units="px", quality=100)

# get what you actually want to plot into a list of lists
initial_ys <- lapply(1:nreps, function(n) {      # loop over replicates
  lapply(replist[[n]][[2]], function(x) {          # loop over timepoints
    nrow(x)})                                      # count the matrix rows = number of individuals at timepoint
  })

#for dataframe, need dimensions to match. Fill in zeros for lists with fewer generations to acheive.
even_ys <- lapply(1:nreps, function(r) {
  even_list = list()
  for (i in 1:maxsamples) {
    if (i%in% 1:length(initial_ys[[r]])) {
      #print(sprintf("Data exists for rep # %s timepoint %s, use 0 for number of individuals", r,i ))
      even_list[i] <- initial_ys[[r]][[i]]
    }
   else {
      #print(sprintf("Data does not exist for rep # %s timepoint %s ", r,i ))
      even_list[i] <- 0
    }}
    return(even_list)
    })

#calculate the mean value for the 20 replicates
meanlist = list()
for (i in 1:maxsamples) { #loop over time
  values = list() #make an empty list to hold the values for all replicates at this one timepoint
  for (j in 1:nreps) { #loop over replicates
    values[j] <- even_ys[[j]][[i]] #sample all replicates, data, this time point
  meanlist[i] <- mean(unlist(values))
  }}  


dataset_list <- c(list(sample_years), list(meanlist), even_ys)
repnames <-unlist(lapply(1:nreps, function(x) {sprintf("Replicate%s",x)}))
columnnames <- c("Years", "Mean", repnames)
names(dataset_list) <- columnnames

# make data into a melted dataframe object for use by ggplot: list of lists, put together lists of counts over time as columns of a matrix, one for each replicate.
data_matrix<- matrix(unlist(dataset_list), ncol=length(dataset_list), byrow=FALSE)
df <- data.frame(data_matrix)
colnames(df) <- columnnames


melteddata <- melt(df, id="Years")
library(ggplot2)

sizes = c(rep(10, 26), rep(1,(length(dataset_list)-2)*26))

sizes = c(2, rep(1,(length(dataset_list)-2)))


ggplot(melteddata, aes(Years, value, colour = variable))+
    geom_line()+
    theme_bw()

dev.off()
```

#old way! don't run - builds genind without saving - not efficient.
## Start exploring pop gen stats: without sampling
``` {r}
print("loading library")
library(adegenet)
print("library loaded")

# find number of alleles over time - find when that is stable at what mutation rate? 
allele_counts_lol<- lapply(1:nreps, function(r) {
  print(sprintf("on rep %s", r))
  mean_allelle_count_list_over_time<-lapply(1:length(replist[[r]][[2]]), function(t){
    print(sprintf("on time %s", t))
    print("building dataframe...")
    rland.df <- data.frame(replist[[r]][[2]][t]) #look at one rland$individuals (rep2, data, timepoint 26)
    print("dataframe built")
    alleles <- seq(7,length(rland.df)-1,2) #sample allele data. start with 7, 29 = length of dataframe-1,
    # Combine alleles per locus to format for what genind wants
    print("reformatting alleles...")
    rland.2genind <- sapply(alleles, function(x){paste(rland.df[,x], rland.df[,x+1], sep = "-")})
    print("alleles formated")
    #label populations
    print("labelling_populations")
    rland.df$pop <- "pop3"
    rland.df$pop[rland.df$X1 < 8] <- "pop2"
    rland.df$pop[rland.df$X1 < 4] <- "pop1"
    print("populations labelled")
    
    #Make a genind object
    print("building genind")
    rland.genind <- df2genind(rland.2genind, pop = rland.df$pop, type = "codom", sep="-")
    print("genind created")
    #how many alleles per locus (average over all populations at a given timepoint)
    print(sprintf("Mean alleles in this rep at this time: %s",mean(rland.genind$loc.n.all) ))
    return(mean(rland.genind$loc.n.all))}) #mean number of alleles per locus
    return(mean_allelle_count_list_over_time)})
```


#use object created above to make another plot.
```{r}
#plot allele counts list of lists

library("reshape2")
jpeg("/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_8_allele_count.jpg", width=600, height=400, units="px", quality=100)

#for plotting dataframe, need dimensions to match. Fill in zeros for lists with fewer generations to acheive.
even_ys <- lapply(1:nreps, function(r) {
  even_list = list()
  for (i in 1:maxsamples) {
    if (i%in% 1:length(allele_counts_lol[[r]])) {
      #print(sprintf("Data exists for rep # %s timepoint %s, use 0 for number of individuals", r,i ))
      even_list[i] <- allele_counts_lol[[r]][[i]]
    }
   else {
      #print(sprintf("Data does not exist for rep # %s timepoint %s ", r,i ))
      even_list[i] <- 0
    }}
    return(even_list)
    })

#calculate the mean value for the 20 replicates
meanlist = list()
for (i in 1:maxsamples) { #loop over time
  values = list() #make an empty list to hold the values for all replicates at this one timepoint
  for (j in 1:nreps) { #loop over replicates
    values[j] <- even_ys[[j]][[i]] #sample all replicates, data, this time point
  meanlist[i] <- mean(unlist(values))
  }}  


dataset_list <- c(list(sample_years), list(meanlist), even_ys)
repnames <-unlist(lapply(1:nreps, function(x) {sprintf("Replicate%s",x)}))
columnnames <- c("Years", "Mean", repnames)
names(dataset_list) <- columnnames

# make data into a melted dataframe object for use by ggplot: list of lists, put together lists of counts over time as columns of a matrix, one for each replicate.
data_matrix<- matrix(unlist(dataset_list), ncol=length(dataset_list), byrow=FALSE)
df <- data.frame(data_matrix)
colnames(df) <- columnnames


melteddata <- melt(df, id="Years")
library(ggplot2)

ggplot(melteddata, aes(Years, value, colour = variable))+
    geom_line()+
    ylab("Mean number of alleles per locus across all pops")
    theme_bw()

dev.off()



```

## Try a new structure: parse data into a genind format (only once - it's a slow step), then gather what's needed from that one variable at a time. This one without sampling.
```{r}
library(adegenet)

genind_lol<- lapply(1:nreps, function(r) { #loop over reps
  print(sprintf("on rep %s", r)) #track progress
  one_reps_geninds_over_time<-lapply(1:length(replist[[r]][[2]]), function(t){#loop over timepoints
    print(sprintf("on time %s", t)) #track progress
    rland.df <- data.frame(replist[[r]][[2]][t]) #df from one rland$individuals 
    alleles <- seq(7,length(rland.df)-1,2) # Combine alleles per locus to format for what genind wants
    rland.2genind <- sapply(alleles, function(x){paste(rland.df[,x], rland.df[,x+1], sep = "-")})
    #label populations
    rland.df$pop <- "pop3"
    rland.df$pop[rland.df$X1 < 8] <- "pop2"
    rland.df$pop[rland.df$X1 < 4] <- "pop1"
    #Make a genind object
    print("building genind")
    rland.genind <- df2genind(rland.2genind, pop = rland.df$pop, type = "codom", sep="-")
    return(rland.genind)}) #mean number of alleles per locus
    return(one_reps_geninds_over_time)})

```

# Working off genind list, get mean alleles per locus again.
```{r}
allele_counts_lol<- lapply(1:nreps, function(r) {
  mean_allelle_count_list_over_time<-lapply(1:length(genind_lol[[r]]), function(t){
    genind <- genind_lol[[r]][[t]]
    return(mean(genind$loc.n.all))}) #mean number of alleles per locus
    return(mean_allelle_count_list_over_time)})

```





## Correction - pop gen stats too slow to run on full datasets. Have to sample at this point. Move sampling here. 
## Then calculate popgen stats
```{r}
library(strataG)
library(adegenet)

overall_test_lol<- lapply(1:nreps, function(r) {
  ot_list_one_rep_over_time<-lapply(1:length(genind_lol[[r]]), function(t){
    print(sprintf("on replicate %s, timepoint %s", r, t))
    genind <- genind_lol[[r]][[t]]
    rland.gtypes <- genind2gtypes(genind)
    ot <- overallTest(rland.gtypes)
    return(ot)}) #mean number of alleles per locus
    return(ot_list_one_rep_over_time)})


#Parse that for overall Fst
overall_Fst_lol<- lapply(1:nreps, function(r) {
  Fst_list_one_rep_over_time<-lapply(1:length(overall_test_lol[[r]]), function(t){
    ot <- overall_test_lol[[r]][[t]]
    Fst <- ot$result
    return(Fst)}) #mean number of alleles per locus
    return(Fst_list_one_rep_over_time)})

#Accessing pieces of this will come next.
#print(ot$result)
#print(ot$result[1,])
#print(ot$result[3,1])
#print(ot$result[3,2])
#print(ot$strata.freq)

pairwise_test_lol<- lapply(1:nreps, function(r) {
  pw_list_one_rep_over_time<-lapply(1:length(genind_lol[[r]]), function(t){
    print(sprintf("on replicate %s, timepoint %s", r, t))
    genind <- genind_lol[[r]][[t]]
    rland.gtypes <- genind2gtypes(genind)
    pw <- pairwiseTest(rland.gtypes)
    return(pw)}) #mean number of alleles per locus
    return(pw_list_one_rep_over_time)})


#Accessing
#pw[1]$result
#pws[1]$result[1,10]


```



## Convert the plot into a function - many things will need to be visulized this way.
```{r}
library("reshape2")
plot_list_of_lists() <- function(input, filename, option, yname){
jpeg(filename, width=600, height=400, units="px", quality=100)
#for plotting dataframe, need dimensions to match (initial data won't). Fill in zeros or NaNs for lists with fewer generations to acheive.
  even_ys <- lapply(1:nreps, function(r) {
  even_list = list()
  for (i in 1:maxsamples) {
    if (i%in% 1:length(input[[r]])) {
      #print(sprintf("Data exists for rep # %s timepoint %s, use 0 for number of individuals", r,i ))
      even_list[i] <- input[[r]][[i]]
    }
   else {
      #print(sprintf("Data does not exist for rep # %s timepoint %s ", r,i ))
      even_list[i] <- option
    }}
    return(even_list)
    })

#calculate the mean value for the replicates
meanlist = list()
for (i in 1:maxsamples) { #loop over time
  values = list() #make an empty list to hold the values for all replicates at this one timepoint
  for (j in 1:nreps) { #loop over replicates
    values[j] <- even_ys[[j]][[i]] #sample all replicates, data, this time point
  meanlist[i] <- mean(unlist(values))
  }}  

dataset_list <- c(list(sample_years), list(meanlist), even_ys)
repnames <-unlist(lapply(1:nreps, function(x) {sprintf("Replicate%s",x)}))
columnnames <- c("Years", "Mean", repnames)
names(dataset_list) <- columnnames

# make data into a melted dataframe object for use by ggplot: list of lists, put together lists of counts over time as columns of a matrix, one for each replicate.
data_matrix<- matrix(unlist(dataset_list), ncol=length(dataset_list), byrow=FALSE)
df <- data.frame(data_matrix)
colnames(df) <- columnnames

melteddata <- melt(df, id="Years")
library(ggplot2)

ggplot(melteddata, aes(Years, value, colour = variable))+
    geom_line()+
    ylab(yname)
    theme_bw()

dev.off()
}

#test:
input_lol <- allele_counts_lol
filepath_to_save <- "/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_8_allele_count.jpg"
option <- 0 #or NaN
y_name <- "Mean allele counts per locus"

plot_list_of_lists(input_lol,filepath_to_save,option,y_name)

```




















#Pop stats with sampling.
```{r}

library(adegenet)
pops <- seq(0,4,2) #get just the flowering plants
rland.df <- data.frame(replist[[2]][[2]][26]) #look at one rland$individuals (rep2, data, timepoint 26)
alleles <- seq(7,length(rland.df)-1,2) #sample allele data. always start with 7, 29 = length of dataframe-1,

# add mapping of columns to what is in them 
# Combine alleles per locus to format for what genind wants
rland.2genind <- sapply(alleles, function(x){
  paste(rland.df[,x], rland.df[,x+1], sep = "-")
})

#label populations
rland.df$pop <- "pop3"
rland.df$pop[rland.df$X1 < 8] <- "pop2"
rland.df$pop[rland.df$X1 < 4] <- "pop1"

#sample from 3, 7, and 11s
table(rland.df$X1)

frompops <- c(3,7,11)

#Sample at random, 30 individuals per population.
rows <- lapply(frompops, function(x){
  n_pop <- length(rland.df$X1[rland.df$X1 == x])
  if(n_pop > 0){
    if(n_pop >= 30){
      sample(which(rland.df$X1 == x),30)
}   else {
      sample(which(rland.df$X1 == x),n_pop)
    }
  }
}
  )

rows <- do.call(c,rows)

samp.rland.2genind <- rland.2genind[rows,]
samp.rland.df <- rland.df[rows,]

library(strataG)

rland.genind <- df2genind(samp.rland.2genind, pop = samp.rland.df$pop, type = "codom", sep="-")
class(rland.genind)
rland.genind

# find number of alleles over time - find when that is stable at what mutation rate? 
#how many alleles per locus (average)
mean(rland.genind$loc.n.all)
#or the full list
rland.genind$loc.n.all
#


rland.gtypes <- genind2gtypes(rland.genind)

ot <- overallTest(rland.gtypes)
print(ot$result)
#print(ot$result[1,])

#print(ot$result[3,1])
#print(ot$result[3,2])


#data.class(ot$result)
#print(ot$strata.freq)


pws <- pairwiseTest(rland.gtypes)
#print("And try to sample")
#print(data.class(pws[1]$result))
pws[1]$result
#pws[1]$result[1,10]


```


## Plot the number of alleles
```{r}
jpeg("/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_8_allele_count.jpg", width=600, height=400, units="px")


dev.off()

```





## Structure of rland$individuals

From "Creating Landscapes" vignette, what the columns in the rland$individuals table mean
"The first six columns are always present and the last five shown here are a product of the choice of number of loci. The first column contains the individuals subpopulation and lifecycle stage (subpopulation = floor(x/rland\$intparam\$stages), lifecycle stage = x mod rland\$intparam\$stages). The second column is currently unused, always 0. The third column contains the generation in which the individual was born or created. The next three contain numerical ids for the individual, its mother and its father. After the first six columns the indivduals genetic code begins. The loci are shown in order with 2 columns for diploid loci and 1 column for haploid loci. The value of these columns represent the allele index of the allele the individual carries."






## Early drafts of other processing step


table of subpopulations
#```{r}


plot(replist[[1]][[1]], nrow(replist[[1]][[2]]))



numind <- sapply(rland.list[[2]], function(x){
  nrow(x)
})
subpops <- lapply(rland.list[[2]], function(x){
  table(x[,1])
})

subpops <- do.call(rbind, subpops)
head(subpops)

subpops <- data.frame(subpops)
subpops1<- melt(subpops, id.vars = rownames(subpops), value.name = "SubClass")

ggplot(subpops, aes())

#```



# Population Genetic statistics
#```{r}
library(adegenet)

pops <- seq(0,4,2)

rland.df <- as.data.frame(replist[[1]][[2]][1])

alleles <- seq(7,29,2)
rland.2genind <- sapply(alleles, function(x){
  paste(rland.df[,x], rland.df[,x+1], sep = "-")
})


rland.df$pop[rland.df$V1 == 0 | rland.df$V1 == 1 ] <- "pop1"
rland.df$pop[rland.df$V1 == 2 | rland.df$V1 == 3 ] <- "pop2"
rland.df$pop[rland.df$V1 == 4 | rland.df$V1 == 5 ] <- "pop3"

# following doesn't work yet!
rland.df$pop <- sapply(pops, function(x){
  pop <- if(rland.df$V1 == x | rland.df$V1 == x+1){
    paste("pop",x,sep="")
  }
  pop
})

str(rland.df)

rland.df2genind <- data.frame(rland.df[,c(3,31)], rland.2genind)

library(strataG)

head(rland.df2genind)

rland.genind <- df2genind(rland.df2genind[,-(1:2)], pop = rland.df2genind$pop, type = "codom", sep="-")
class(rland.genind)
rland.genind

rland.gtypes <- genind2gtypes(rland.genind)

overallTest(rland.gtypes)
pairwiseTest(rland.gtypes)



#```




