---
title: "Rmetasim_processing_saved_data_V6"
author: "Jessie Berta-Thompson"
date: "February 10, 2017"
output:
  html_document: default
  pdf_document: default
---
# not in today: track hours - 12:40-1:50 (log 1 hour)
(but couldn't help taking one more shot at this annoying bug)

# immediate purpose of this version: deal with cutoff in sampling step - full of errors!

# short term steps coming up
## incorporate population genetics
## implement sampling versions in same framework
## perform a test of different sampling sizes (then you're going to need stats to test detectability of features with different n)

## using V6 2000 plants + seeds (2% plants total) + 5000 generations + 10 replicates test run while developing this script

## Structure of Rmetasim output files that this script parses
Output datafile from each Rmetasim run contains a 'replist' list of lists with generations sampled and rland$individuals for each replicate, the structure of which is:
[[[rep1 generations sampled list][rep1 rland$individuals list of matrices]] [[rep2 generations sampled list][rep2 rland$individuals list of matrices]]...]
Replist is a list containing r objects, where r is the number of replicates, with a dataset for each replicate.
Each of the replicate objects is a list containing 2 objects, a list containing the year for t generations sampled and a list containing rland$individuals matrices with genetic data for each t generation sampled.
Sampling from nested lists:
The first replicate's (1) individual list (2) containing rland$individuals for all time points: replist[[1]][[2]]
The first replicate's (1) list of generation times (1) that correspond to samples in individual list: replist[[1]][[1]]
The first replicate's (1) individual's data (2) for the third timepoint (3) (a single rland$individuals object): [[1]][[2]][3]

## Processing rmetasim output
```{r}
#clear environment before starting
rm(list =ls())
#load desired dataset
load("/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen.Rdata") 

#usual path on lab computer example
#load("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Rmetasim/replist_1_19_17_V6_seed_test_5000gen.Rdata")
```


## Extract some useful values about the dataset from its dimensions
```{r}
#How many replicates does this dataset contain?
nreps = length(replist)
print(sprintf("This dataset contains %s replicate simulations.", nreps))

#How many time samples were collected for each replicate? How many time samples maximum? (a given replicate can have less than the input length of run when extinctions occur)
nsamples_list <- list()
for (i in 1:length(replist)){
  nsamples <-length(replist[[i]][[1]])
  nsamples_list[i] <- nsamples}

maxsamples <- max(unlist(nsamples_list))
print(sprintf("At most, %s time samples were collected.", maxsamples))

#How many years of simulation does this represent?
nyears_list <- list()
for (i in 1:length(replist)){
  nyears <-replist[[i]][[1]][length(replist[[i]][[1]])]
  nyears_list[i] <- nyears}

maxyears <- max(unlist(nyears_list))
print(sprintf("At most, this simulation was run for %s years.", maxyears))

#what years were sampled (for the longest)?
index_of_longest <- which.max(unlist(nyears_list))
sample_years <- replist[[index_of_longest]][[1]]
print("These years were sampled:")
print(unlist(sample_years))

```



## Plotting function - many things will need to be visulized this way.
```{r}
library("reshape2")
library("ggplot2")

plot_list_of_lists <- function(input, file_name, option, yname){
#for plotting dataframe, need dimensions to match (initial data won't). Fill in zeros or NaNs for lists with fewer generations to acheive.
  even_ys <- lapply(1:nreps, function(r) {
  even_list = list()
  for (i in 1:maxsamples) {
    if (i%in% 1:length(input[[r]])) {
      #print(sprintf("Data exists for rep # %s timepoint %s, use 0 for number of individuals", r,i ))
      even_list[i] <- input[[r]][[i]]}
   else {
      #print(sprintf("Data does not exist for rep # %s timepoint %s ", r,i ))
      even_list[i] <- option
    }}
    return(even_list)})

#calculate the mean value for the replicates
meanlist = list()
for (i in 1:maxsamples) { #loop over time
  values = list() #make an empty list to hold the values for all replicates at this one timepoint
  for (j in 1:nreps) { #loop over replicates
    values[j] <- even_ys[[j]][[i]] #sample all replicates, data, this time point
  meanlist[i] <- mean(unlist(values))}}  

dataset_list <- c(list(sample_years), list(meanlist), even_ys)
repnames <-unlist(lapply(1:nreps, function(x) {sprintf("Replicate%s",x)}))
columnnames <- c("Years", "Mean", repnames)
names(dataset_list) <- columnnames

# make data into a melted dataframe object for use by ggplot: list of lists, put together lists of counts over time as columns of a matrix, one for each replicate.
data_matrix<- matrix(unlist(dataset_list), ncol=length(dataset_list), byrow=FALSE)
df <- data.frame(data_matrix)
colnames(df) <- columnnames

melteddata <- melt(df, id="Years")

p<- ggplot(melteddata, aes(Years, value, colour = variable))+
    geom_line()+
    ylab(yname)
    theme_bw()
ggsave(filename=file_name, plot=p, width = 6, height = 4, units = "in")
}
```


Gather first data set: simple counts of individuals.
```{r}
#build a lists of lists (what plotting function accepts), for the total number of individuals at each timepoint.
total_rland_ind_counts_lol <- lapply(1:nreps, function(n) {      # loop over replicates
  lapply(replist[[n]][[2]], function(x) {nrow(x)})})           # loop over timepoints, count the matrix rows = number of individuals at timepoint 
```


Test plotting function on that count data
```{r}
#test:
input_lol <- total_rland_ind_counts_lol
filepath_to_save <- "/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_9_total_ind_counts.jpg"
option <- 0 #or NaN (average will be calculated - decide if 0s or ignore more approriate for post-extinction time points)
y_name <- "Total individuals"

plot_list_of_lists(input=input_lol,file_name=filepath_to_save,option=option,yname=y_name)

#input, file_name, option, yname can also be entered in order
```


How many populations remaining? (3, 2, 1 or 0)
```{r}
#build a lists of lists (what plotting function accepts), for the total number of individuals at each timepoint.
npops_lol <- lapply(1:nreps, function(r) {      # loop over replicates
  lapply(replist[[r]][[2]], function(x) { # loop over timepoints
    rland.df <- data.frame(x) 
    nsubpops<- length(unique(rland.df[,1]))
    if (nsubpops %in% c(1, 2, 3, 4)) {
      npops <- 1
      }
    else if (nsubpops %in% c(5, 6, 7, 8)){
      npops <- 2
      }
    else if (nsubpops %in% c(9, 10, 11, 12)){
      npops <- 3
      }
    else if (nsubpops == 0) {
      npops <- 0
    }
    else {
      print("something weird")
    }
    print(sprintf("Rep %s, one timepoint, npops %s", r, npops))
    return(npops)
    })
    })    
```


Plot this number of populations list.
```{r}
#test:
input_lol <- npops_lol
filepath_to_save <- "/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_9_number_of_pops.jpg"
option <- 0 #or NaN (average will be calculated - decide if 0s or ignore more approriate for post-extinction time points)
y_name <- "Number of populations"

plot_list_of_lists(input=input_lol,file_name=filepath_to_save,option=option,yname=y_name)

#input, file_name, option, yname can also be entered in order
```


## Parse data into a genind format without sampling (only once - it's a slow step), then gather what's needed from that one variable at a time.
```{r}
library(adegenet)
#*****if already saved, just run this line instead. but even just loading these can be slow.
load("/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_genind_lol.Rdata")

genind_lol<- lapply(1:nreps, function(r) { #loop over reps
  print(sprintf("on rep %s", r)) #track progress
  one_reps_geninds_over_time<-lapply(1:length(replist[[r]][[2]]), function(t){#loop over timepoints
    print(sprintf("on time %s", t)) #track progress
    rland.df <- data.frame(replist[[r]][[2]][t]) #df from one rland$individuals 
    alleles <- seq(7,length(rland.df)-1,2) # Combine alleles per locus to format for what genind wants
    rland.2genind <- sapply(alleles, function(x){paste(rland.df[,x], rland.df[,x+1], sep = "-")})
    #label populations
    rland.df$pop <- "pop3"
    rland.df$pop[rland.df$X1 < 8] <- "pop2"
    rland.df$pop[rland.df$X1 < 4] <- "pop1"
    #Make a genind object
    print("building genind")
    rland.genind <- df2genind(rland.2genind, pop = rland.df$pop, type = "codom", sep="-")
    return(rland.genind)})
    return(one_reps_geninds_over_time)})

#because this step is slow, save the list of lists of geninds
save(genind_lol,file="/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_genind_lol.Rdata")

```

# Working off genind list, get mean alleles per locus again.
```{r}
allele_counts_lol<- lapply(1:nreps, function(r) {
  mean_allelle_count_list_over_time<-lapply(1:length(genind_lol[[r]]), function(t){
    genind <- genind_lol[[r]][[t]]
    return(mean(genind@loc.n.all))}) #mean number of alleles per locus
    return(mean_allelle_count_list_over_time)})

```

Plot this mean alleles per locus list.
```{r}
#test:
input_lol <- allele_counts_lol
filepath_to_save <- "/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_9_allele_counts.jpg"
option <- 0 #or NaN (average will be calculated - decide if 0s or ignore more approriate for post-extinction time points)
y_name <- "Mean number of alleles per locus"

plot_list_of_lists(input=input_lol,file_name=filepath_to_save,option=option,yname=y_name)

#input, file_name, option, yname can also be entered in order
```

## Full pop gen stats too slow to run on full datasets. Have to sample at this point. 
#Pop stats with sampling.
```{r}
library(adegenet) # multivariate genetic data analysis
library(strataG)  # summaries and population structure analyses of genetic data

sampled_genind_lol<- lapply(1:nreps, function(r) {# loop over replicates
  sampled_genind_list_over_time<-lapply(1:length(replist[[r]][[2]]), function(t){# loop over matrices 
    print(sprintf("Working in replicate %s, time %s", r, t))
    rland.df <- data.frame(replist[[r]][[2]][t]) #convert one individuals matrix to dataframe format
    #reformat allele data (pairs by locus)
    alleles <- seq(7,length(rland.df)-1,2) # in rland$individuals, alleles start at 7, go to end, pair at each locus side by side
    rland.2genind <- sapply(alleles, function(x){
      paste(rland.df[,x], rland.df[,x+1], sep = "-")}) # Combine alleles per locus with '-' for genind
    #convert to data frame to match the other one.
    rland.2genind <- data.frame(rland.2genind)
    ### For debugging: 
    print("Dimensions of full rland.df (rxc):")
    print(dim(rland.df))
    print("Dimensions of full rland.2genind (rxc):")    
    print(dim(rland.2genind))

    #label with populations [subpopulations 0, 1, 2, 3 = pop1, 4, 5, 6, 7 = pop2, 8, 9, 10, 11 = pop3]
    rland.df$pop <- "pop3"
    rland.df$pop[rland.df$X1 < 8] <- "pop2"
    rland.df$pop[rland.df$X1 < 4] <- "pop1"
    
    #sample just flowering plants from each population: 3, 7, and 11s for 3 populations with 4 age classes in each.
    flowers <- c(3,7,11)
    #sample at random, n individuals per population.
    n <- 30 #number of samples to take. will sample exactly that many - doesn't bother if fewer than 30 flowers.
    sample_rows <- lapply(flowers, function(x){ #loop through flower pops to create sample list.
      n_pop <- length(rland.df$X1[rland.df$X1 == x]) #count the number of instances of given population x. X1 contains pop id (3, 7, 11)
      print(sprintf("population %s contains %s plants", x, n_pop))
      if (n_pop>=n) {
        print(sprintf("enough plants in population %s to sample target %s individuals!",x,n ))
        rowstotake <- sample(which(rland.df$X1 == x),n) #sample at random 30 items from the indices of rows in correct pop
      }
      else{
        print(sprintf("not enough plants in population %s to sample target %s individuals: returning NULL",x,n ))
        rowstotake <- NULL
      }
      return(rowstotake)
      })    
     print(sample_rows) #this is a list of lists
    rows <- do.call(c,sample_rows) #this collapses list of lists into a single vector.
    #this doesn't have special case handling when all three are below 30 - ignore for now, but could be a problem.
    print(rows)
    #apply sampling to two data frames
    samp.rland.2genind <- rland.2genind[rows,] # sample genind-style data (just alleles) for the sample rows, all columns
    samp.rland.df <- rland.df[rows,]

    ### For debugging: 
    print("Dimensions of samp.rland.2genind (rxc):")
    print(dim(samp.rland.2genind))
    print("Dimensions of samp.rland.df (rxc):")    
    print(dim(samp.rland.df))
    
    #convert to genind for downstream analyses
    samp.rland.genind <- df2genind(samp.rland.2genind, pop = samp.rland.df$pop, type = "codom", sep="-") #error occurs here - size mismatch
       return(samp.rland.genind)})
    return(sampled_genind_list_over_time)})
```



#repeat the loop and logic structure of sampling to build a report on sampling success (when was there enough), without the work
```{r}
sample_report_lol<- lapply(1:nreps, function(r) {# loop over replicates
  sample_report_list_over_time<-lapply(1:length(replist[[r]][[2]]), function(t){# loop over matrices 
    print(sprintf("Working in replicate %s, time %s", r, t))
    rland.df <- data.frame(x) #convert one individuals matrix to dataframe format
    #label with populations [subpopulations 0, 1, 2, 3 = pop1, 4, 5, 6, 7 = pop2, 8, 9, 10, 11 = pop3]
    rland.df$pop <- "pop3"
    rland.df$pop[rland.df$X1 < 8] <- "pop2"
    rland.df$pop[rland.df$X1 < 4] <- "pop1"
    #sample just flowering plants from each population: 3, 7, and 11s for 3 populations with 4 age classes in each.
    flowers <- c(3,7,11)
    #sample at random, n individuals per population.
    n <- 30 #number of samples to take. will sample exactly that many - doesn't bother if fewer than 30 flowers.
    sample_success_vector <- sapply(flowers, function(x){ #loop through flower pops to create sample list.
      n_pop <- length(rland.df$X1[rland.df$X1 == x]) #count the number of instances of given population x. X1 contains pop id (3, 7, 11)
      if (n_pop>=n) {
        outcome <-1
      }
      else{
        outcome<-0
      }
      return(outcome)
      })    
         return(sample_success_vector)})
    return(sample_report_list_over_time)})

# ask michelle for ideas on how to visualize this nicely. might need to find a way to handle cases where it dips below 30 then comes back.
# not sure if pop gen stuff works below 3 populations (errors downstream) - need to think. We aren't modelling the way things are now if we lose any populations, so requiring all 3 might be another way to go. 

```



# Working off sampled genind list, calculate overall test pop gen stats
```{r}
sampled_overall_test_lol<- lapply(1:nreps, function(r) {
  sampled_overall_list_over_time<-lapply(1:length(sampled_genind_lol[[r]]), function(t){
    genind <- sampled_genind_lol[[r]][[t]]
    rland.gtypes <- genind2gtypes(genind)
    print(sprintf("Calculating overall test for rep %s timepoint %s ", r, t))
    ot <- overallTest(rland.gtypes)
    return(ot)}) #mean number of alleles per locus
    return(sampled_overall_list_over_time)})

```

      print(ot$result)
#print(ot$result[1,])
#print(ot$result[3,1])
#print(ot$result[3,2])

#data.class(ot$result)
#print(ot$strata.freq)
pws <- pairwiseTest(rland.gtypes)
#print("And try to sample")
#print(data.class(pws[1]$result))
pws[1]$result
#pws[1]$result[1,10]



```{r}
library(strataG)
library(adegenet)

overall_test_lol<- lapply(1:nreps, function(r) {
  ot_list_one_rep_over_time<-lapply(1:length(genind_lol[[r]]), function(t){
    print(sprintf("on replicate %s, timepoint %s", r, t))
    genind <- genind_lol[[r]][[t]]
    rland.gtypes <- genind2gtypes(genind)
    ot <- overallTest(rland.gtypes)
    return(ot)}) #mean number of alleles per locus
    return(ot_list_one_rep_over_time)})


#Parse that for overall Fst
overall_Fst_lol<- lapply(1:nreps, function(r) {
  Fst_list_one_rep_over_time<-lapply(1:length(overall_test_lol[[r]]), function(t){
    ot <- overall_test_lol[[r]][[t]]
    Fst <- ot$result
    return(Fst)}) #mean number of alleles per locus
    return(Fst_list_one_rep_over_time)})

#Accessing pieces of this will come next.
#print(ot$result)
#print(ot$result[1,])
#print(ot$result[3,1])
#print(ot$result[3,2])
#print(ot$strata.freq)

pairwise_test_lol<- lapply(1:nreps, function(r) {
  pw_list_one_rep_over_time<-lapply(1:length(genind_lol[[r]]), function(t){
    print(sprintf("on replicate %s, timepoint %s", r, t))
    genind <- genind_lol[[r]][[t]]
    rland.gtypes <- genind2gtypes(genind)
    pw <- pairwiseTest(rland.gtypes)
    return(pw)}) #mean number of alleles per locus
    return(pw_list_one_rep_over_time)})


#Accessing
#pw[1]$result
#pws[1]$result[1,10]


```




















## Plot the number of alleles
```{r}
jpeg("/Users/jessie/Documents/botanic/2_3_17_2_8_17_coding/replist_1_19_17_V6_seed_test_5000gen_2_8_allele_count.jpg", width=600, height=400, units="px")


dev.off()

```





## Structure of rland$individuals

From "Creating Landscapes" vignette, what the columns in the rland$individuals table mean
"The first six columns are always present and the last five shown here are a product of the choice of number of loci. The first column contains the individuals subpopulation and lifecycle stage (subpopulation = floor(x/rland\$intparam\$stages), lifecycle stage = x mod rland\$intparam\$stages). The second column is currently unused, always 0. The third column contains the generation in which the individual was born or created. The next three contain numerical ids for the individual, its mother and its father. After the first six columns the indivduals genetic code begins. The loci are shown in order with 2 columns for diploid loci and 1 column for haploid loci. The value of these columns represent the allele index of the allele the individual carries."






## Early drafts of other processing step


table of subpopulations
#```{r}


plot(replist[[1]][[1]], nrow(replist[[1]][[2]]))



numind <- sapply(rland.list[[2]], function(x){
  nrow(x)
})
subpops <- lapply(rland.list[[2]], function(x){
  table(x[,1])
})

subpops <- do.call(rbind, subpops)
head(subpops)

subpops <- data.frame(subpops)
subpops1<- melt(subpops, id.vars = rownames(subpops), value.name = "SubClass")

ggplot(subpops, aes())

#```



# Population Genetic statistics
#```{r}
library(adegenet)

pops <- seq(0,4,2)

rland.df <- as.data.frame(replist[[1]][[2]][1])

alleles <- seq(7,29,2)
rland.2genind <- sapply(alleles, function(x){
  paste(rland.df[,x], rland.df[,x+1], sep = "-")
})


rland.df$pop[rland.df$V1 == 0 | rland.df$V1 == 1 ] <- "pop1"
rland.df$pop[rland.df$V1 == 2 | rland.df$V1 == 3 ] <- "pop2"
rland.df$pop[rland.df$V1 == 4 | rland.df$V1 == 5 ] <- "pop3"

# following doesn't work yet!
rland.df$pop <- sapply(pops, function(x){
  pop <- if(rland.df$V1 == x | rland.df$V1 == x+1){
    paste("pop",x,sep="")
  }
  pop
})

str(rland.df)

rland.df2genind <- data.frame(rland.df[,c(3,31)], rland.2genind)

library(strataG)

head(rland.df2genind)

rland.genind <- df2genind(rland.df2genind[,-(1:2)], pop = rland.df2genind$pop, type = "codom", sep="-")
class(rland.genind)
rland.genind

rland.gtypes <- genind2gtypes(rland.genind)

overallTest(rland.gtypes)
pairwiseTest(rland.gtypes)



#```




