---
title: "Rmetasim_v4"
author: "Michelle DePrenger-Levin & Jessie"
date: "January 12, 2017"
output: html_document
---
Ctrl + Alt + c to run chunk 

To do:

1) implement trycatch to cover weird instances of rmetasim crashing. (i like this explanation)
http://stackoverflow.com/questions/12193779/how-to-write-trycatch-in-r

2) Work up population genetics stats from output

3) Add fourth life stage for seedbank

4) lots of parameter choice research & testing



Running before I leave. parameters to test population stability with much higher carrying capacity.


To start, clear environment and load rmetasim
```{r}
rm(list =ls())
library(rmetasim)
library(dplyr)
```


Set values for integer parameters
```{r}
#the number of different habitats or subpopulations within the landscape - for now, we want 3 Phacelia populations
habitats <- 3
#the number of stages in the life cycle of the organism; also the size of the S, R, and M matricies for local demography - for the roughly biennial Phacelia, we want 3 life stages: seedlings, still juvenile year-old rosettes, flowering plants
stages <- 3 
# 12 microsatellite loci used in genotyping study [0] "the number of different loci currently implemented for the simulation"
locusnum <- 12
# one migration scenario over time [0] "the number of different migration scenerios to occur during the simulation"
numepochs <- 0   
# start at generation 0 [0] "the current generation the simulation has reached" 
currentgen <- 0   
# start at epoch 0 [0] "the current epoch the simulation has reached"
currentepoch <- 0  
# run for at most 100,000 generations; in practice this is set later: .simulate(numit)    "the total number of generations to simulate"
totalgens <- 100000  
# only one within-population demography (life cycle) for all populations [0] "the number of within population demographies"
numdemos <- 0  
# overall landscape limit set to be bigger than populations expected to go [2e+05] "the maximum number of individuals that can exist in the simulation"
maxlandsize <- 5e+05      
```


Set on/off values for parameters that are switches: 4 values of 1 or 0
```{r}
randepoch <- 0 # only one epoch - no need for random transitions [0] "1=choose the next epoch randomly using their individual probabilities, 0=choose the next epoch according to their ordering"     
randdemo <- 0 # only one local demography across the board - no need for random assignments [0] "1=assign demographies at random, 0=assign demographies in order"                                                                
multp <- 1 # multiple paternity on, females can be pollinated by more than one male plant [1] "1=multiple paternity, 0=entire families from a single mating"                                                                   
densdepdemo <- 0 # a single demography will apply regardless of density [0] "a flag to turn density-dependent population regulation on and off"                                
```


Set selfing rate (float parameter)
We don't know selfing rate; Phacelia attracts pollinators (looking for outside pollen), but doesn't have morphology to absolutely prevent selfing (more research required + run various  to assess impact)
```{r}
selfing <- 0.5      # guess with some selfing, some outside pollen      [0] "This value is the selfing rate of the species (between 0 and 1 inclusive) assuming a mixed mating model where selfing occurs S proportion of the time and random mating occurs 1-S proportion of the time"
```

Set Local demographics (life stages) matrices and parameters
Local demography - within-population life cycle dynamics - is composed of survival, reproduction and male contribution matrices, which are combined to advance the population each generation
There can be a different one for each subpopulation (chosen at random from the set or assigned in order)
Local demography S, R, M matrices must be square n x n, where n is the number of stages in each population  
Repeating landscape.new.local.demo with more matrices puts more demographies into the localdem list (e.g. one for each population)

```{r}
#Local Survival
localS <- matrix(c(              0,     0,   0,
                    0.011111111111,     0,   0,
                             0.091,  0.81,   0) , byrow = TRUE, nrow = 3)
#Local Reproduction
localR <- matrix(c(  0,  0,  10,
                     0,  0,  0,
                     0,  0,  0) , byrow = TRUE, nrow = 3)
#Local Male Contribution
localM <- matrix(c(  0,  0,  0,
                     0,  0,  0,
                     0,  0,  1) , byrow = TRUE, nrow = 3)

```

Explanation of demography choices for a slightly complicated biennial
stage 1 = seedlings, new rosettes
stage 2 = juvenile rosettes created from last year's seedlings and waiting until the next spring to flower (an extra non reproductive year)
stage 3 = flowering plants that reproduce then die, created from last years seedlings and year old juvenile rosettes.
no stage is allowed to linger - individuals either move to another stage or die every year.
values set to give a stable population that is 90% seedlings, 1% juveniles, 9% flowering plants

Survival
reading down each column (transitions go from column to row)
No seedlings remain seedlings, ~1% become juveniles, ~9% become flowering plants next year [~90% die]
No juveniles become seedlings, no juveniles remain juveniles, 75% of juveniles survive to become flowering plants next year [25% die]
No mature flowering plants become seedlings or juveniles and none survive another year [100% die]

Reproduction
Maternal production of offspring
Our choice means each flowering plant contributes 10 seedlings to the next generation, on average (drawn from poisson about 10)

Male contribution - Probability that pollen comes from a particular stage class.
All pollen comes from flowering plants
#****** Question ******** should this value be 1 or 1-contents of migration matrices? how is this combined with the epoch migration matrices? 100% percent of pollen comes from flowering plants, but we really want 90% from within population, 10% from other populations.



Between-population demographics (migration)
Repeat command to have more epochs
Construct matrices [number of habitats x number of stages] x [number of habitats x number of stages] describing movement between all demographic subpopulations of all populations.
No plants travel (survival matrix all zero. Seeds might travel (Reproduction matrix), but zero for now.
Pollen travels. A small amount of male contribution comes from other populations (could use clarification on exactly how this is calculated, from/to)
```{r}
epochS <- matrix(c(rep(0,81)), nrow=9)
epochR <- matrix(c(rep(0,81)), nrow=9)
epochM <- matrix(c(   0,  0,    0,     0,  0,    0,    0,  0,    0,
                  0,  0,    0,     0,  0,    0,    0,  0,    0,
                  0,  0,    0,     0,  0, 0.05,    0,  0, 0.05,
                  
                  0,  0,    0,     0,  0,    0,    0,  0,    0,
                  0,  0,    0,     0,  0,    0,    0,  0,    0,
                  0,  0, 0.05,     0,  0,    0,    0,  0, 0.05,
                  
                  0,  0,    0,     0,  0,    0,    0,  0,    0,
                  0,  0,    0,     0,  0,    0,    0,  0,    0,
                  0,  0, 0.05,     0,  0, 0.05,    0,  0,    0), byrow=TRUE, nrow=9)
```

#After the matrices, the epoch set-up command also accepts other arguments about demography
```{r}
RndChooseProb <- 1            
# random flag not turned on, unimportant, leave default           [1]                 "the probability of randomly choosing this epoch if rland$switchparam$randepoch==1"
StartGen <- 0                
# this epoch will staert at generation 0                          [0]                 "the generation this epoch begins if rland$switchparam$randepoch==0"
Extinct <- c(0,0,0)           
# no extinction in model (could happen if populations shrink)     [c(0,0,0)]          "a vector of values giving the extinction probability for each subpopulation"
Carry <- c(1e+5,1e+5,1e+5)    
# carrying capacities set to higher than expected so not limiting [c(1000,1000,1000)] "a vector of values denoting the maximum individuals for each subpopulation"

Localprob <- 1                
# the probability of choosing the one local demography is 1       [1]                 "the probability of choosing this local demography if rland$switchparam$randdemo==1"
```


Create lists of loci and alleles
Run landscape.new.locus command once for each locus. Set information about locus. 

Each locus also has an object called alleles, which is a list of alleles for the locus, each of which is a list with more information. Allele properties at a locus are also fed into landscape.new.locus to populate initial allele frequencies

One unused property: allelesize <- 50  # no allele size used for strict stepwise model [50] for type==2 DNA base substitution/variable length sequence state, enter custom DNA length here (one for locus, not per allele, so not quite what we want for microsatellite-like evolution)   
```{r}
model <- 1          
# 0 = infinite alleles (alleles represented by integers, every mutation creates a new allele, no back mutation) sometimes used for compound microsatellites with complex patterns of mutation with less likely back mutation.
# 1 = strict step wise (alleles represented by integers, increase or decrease by 1 every mutation - like differences in length for simple sequence repeats - used for microsatellites with repeat regions, co-dominant, high mutation rates) 
# 2 = DNA base substitution (probably JC)
ploidy <- 2             
# diploid                                                                 [1]       "This can be 1 for haploid and 2 for diploid"
mu <- 0.0001  
# per-allele mutation rate 0.0001                                         [0]       "the per-allele mutation rate (not per site) range [0 1] (average # of mutations per allele per generation)"
transmission <- 0       
# biparental inheritance                                                  [1]       "The mode of inheritance. This can be 0 for biparental or 1 for maternal."
numalleles <- 9  # number of alleles at this locus (before mutation). allow there to be n possible states of the microsatellite at this locus.

#set up a frequency list for the n alleles at this locus (initialize with even frequencies of all alleles) (Same length as number of alleles)
freq <- c(rep(1.0/numalleles,numalleles)) 

```
Additional features of locus and allele objects.

statelist <- # the starting alleles themselves - no need to enter, they are generated by program (although you could) for type 0 or 1 this fills in integers [1:numalleles]; for type 2, generates DNA of length allelesize sampling ACTG from uniform distribution
landscape.new.locus the loci parameters and makes an the allele list of length numalleles so that each allele gets the following properties:
aindex index number of allele
birth generation the allele first appeared
prop - frequency of this allele (from frequency list)
state - an integer for loci of type 0 or 1 (infinite allele or strict stepwise mutation), a string representing DNA sequence for loci type 2 (ACTG)
all possible entries: rland <- landscape.new.locus(rland, type=locustype, ploidy=ploidy, mutationrate=mutationrate, transmission=transmission, numalleles=numalleles, allelesize=allelesize,frequencies=frequencylist, states=statelist)
If for some reason we wanted to get into allele construction, code to do so (from landscape.new.locus) would look like this
rland$loci[[locusnum]]$alleles <- makealleles(type, numalleles, allelesize, frequencies, states)


To create individuals, enter a vector of population sizes for each subpopulation, summing to less than carrying capacity for each.
For 3 populations with 3 subpopulations, these 9 categores become known as subpopulations 0-8 in data (in order shown)
```{r}
pop1seedling <- 45000
pop1juv      <- 500
pop1mat      <- 4500

pop2seedling <- 45000
pop2juv      <- 500
pop2mat      <- 4500

pop3seedling <- 45000
pop3juv      <- 500
pop3mat      <- 4500

initial_pop_structure = c(pop1seedling, pop1juv, pop1mat, pop2seedling, pop2juv, pop2mat, pop3seedling, pop3juv, pop3mat)
```


#function to create an Rland!
```{r}

create.land <- function()
    {
### CREATE A LANDSCAPE
###first set up the matrices for local demographies
        S <- localS
        R <- localR
        M <- localM
        #and epochs 
        S.epoch <- epochS
        R.epoch <- epochR
        M.epoch <- epochM

        ##now create the landscape
        landscape.new.empty() %>% 
            landscape.new.intparam(h=habitats,s=stages, cg=currentgen, ce=currentepoch,
                                   totgen=totalgens, maxland=maxlandsize) %>% 
                landscape.new.switchparam(re = randepoch, rd = randdemo, 
                                          mp = multp, dd =densdepdemo) %>%
                    landscape.new.floatparam(s = selfing) %>%
                        landscape.new.local.demo( S, R, M) %>%
                            landscape.new.epoch(S=epochS, R=epochR, M=epochM, 
                                                epochprob = RndChooseProb, startgen = StartGen, 
                                                extinct = Extinct, carry = Carry, 
                                                localprob = Localprob) %>%
          #12 loci
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%
          landscape.new.locus(type=model, ploidy=ploidy, mutationrate=mu,transmission=transmission, numalleles=numalleles, frequencies=freq) %>%

          landscape.new.individuals(initial_pop_structure)
    }
```


#Test create.land function
```{r}
rland <- create.land()

rland$individuals

```

Run the code
```{r}
#How many times to replicate simulation?
rep <- 20 #set reps of same run
t00 = Sys.time()
replist <- lapply(1:rep, function(r){
  #Which generations would you like to keep?
  maxgen <- 10000 #change as needed
  howmany <- 100 # how many times do you want to keep data
  samplegens <- floor(seq(1,maxgen, maxgen/howmany))
  
  #make a lanscape
  rland <- create.land()
  
  #set up a datastructure to hold datasets saved for analysis
  individual_list = list() #sampling generations plus T0 and Tf
  individual_list[[1]] <- rland$individuals
  generation_list = list()
  generation_list[[1]] <- 0
  samplingcounter = 2
  
  for (gen in 1:maxgen) {
    t0 = Sys.time()
    rland <- landscape.extinct(rland)
    rland <- landscape.survive(rland)
    rland <- landscape.reproduce(rland)
    rland <- landscape.carry(rland)
    rland <- landscape.advance(rland)
    
    if (gen %in% samplegens) {
      individual_list[[samplingcounter]] <- rland$individuals
      generation_list[[samplingcounter]] <- rland$intparam$currentgen
      samplingcounter <- samplingcounter + 1
      }
    tf = Sys.time()
    print(sprintf("Finished replicate %s of %s, generation %s of %s in %s sec. %s min since start.",r, rep, rland$intparam$currentgen, maxgen, format(as.numeric(tf-t0,units="secs"), digits=3), format(as.numeric(tf-t00,units="mins"), digits=3)))
print(table(rland$individuals[,1]))
    }

#top off the data files with the final values
individual_list[[samplingcounter]] <- rland$individuals
generation_list[[samplingcounter]] <- rland$intparam$currentgen

#create a single object to output
output <- list(generation_list, individual_list)






})

# one ran on the 12th lost all individuals

save(replist, file =
"Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Rmetasim/replist_1_13_17_V4_high_carrying_capacity_10000gen.Rdata")
```


From "Creating Landscapes" vignette, what the columns in the rland$individuals table mean
"The first six columns are always present and the last five shown here are a product of the choice of number of loci. The first column contains the individuals subpopulation and lifecycle stage (subpopulation = floor(x/rland\$intparam\$stages), lifecycle stage = x mod rland\$intparam\$stages). The second column is currently unused, always 0. The third column contains the generation in which the individual was born or created. The next three contain numerical ids for the individual, its mother and its father. After the first six columns the indivduals genetic code begins. The loci are shown in order with 2 columns for diploid loci and 1 column for haploid loci. The value of these columns represent the allele index of the allele the individual carries."


#Plot some simple features of the results 


```{r}


x1 <- replist[[1]][[1]] #list with values for generations sampled in replicate 1
y1 <- lapply(replist[[1]][[2]], function(x) {nrow(x)})

plot(x1,y1,ylim=c(0,30000))
lines(x1,y1)

for (i in 2:rep) {
xs <- replist[[i]][[1]] #list with values for generations sampled in replicate 1
ys <- lapply(replist[[i]][[2]], function(x) {nrow(x)})
 # list with matrices; count length of each matrix
points(xs,ys)
lines(xs,ys)

}


```


```{r}


```





table of subpopulations
```{r}






plot(replist[[1]][[1]], nrow(replist[[1]][[2]]))



numind <- sapply(rland.list[[2]], function(x){
  nrow(x)
})
subpops <- lapply(rland.list[[2]], function(x){
  table(x[,1])
})

subpops <- do.call(rbind, subpops)
head(subpops)

subpops <- data.frame(subpops)
subpops1<- melt(subpops, id.vars = rownames(subpops), value.name = "SubClass")

ggplot(subpops, aes())

```



# Population Genetic statistics
```{r}
library(adegenet)

rland.df <- as.data.frame(rland.df)

pops <- seq(0,4,2)

alleles <- seq(7,29,2)
rland.2genind <- sapply(alleles, function(x){
  paste(rland.df[,x], rland.df[,x+1], sep = "-")
})


rland.df$pop[rland.df$V1 == 0 | rland.df$V1 == 1 ] <- "pop1"
rland.df$pop[rland.df$V1 == 2 | rland.df$V1 == 3 ] <- "pop2"
rland.df$pop[rland.df$V1 == 4 | rland.df$V1 == 5 ] <- "pop3"

# following doesn't work yet!
rland.df$pop <- sapply(pops, function(x){
  pop <- if(rland.df$V1 == x | rland.df$V1 == x+1){
    paste("pop",x,sep="")
  }
  pop
})

str(rland.df)

rland.df2genind <- data.frame(rland.df[,c(3,31)], rland.2genind)

library(strataG)

head(rland.df2genind)

rland.genind <- df2genind(rland.df2genind[,-(1:2)], pop = rland.df2genind$pop, type = "codom", sep="-")
class(rland.genind)
rland.genind

rland.gtypes <- genind2gtypes(rland.genind)

overallTest(rland.gtypes)
pairwiseTest(rland.gtypes)



```






