---
title: "Rmetasim_processing_saved_data_V1"
author: "Jessie Berta-Thompson"
date: "January 19, 2017"
output:
  html_document: default
  pdf_document: default
---
## Processing rmetasim output

Clean the environment and load the datafile containing a 'replist' with generations sampled and rland$individuals
```{r}
rm(list =ls())
load("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Rmetasim/replist_1_19_17_V6_seed_test_5000gen.Rdata")
#load("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Rmetasim/replist_1_18_17_V5_stability_test_10000gen.Rdata")

```

Extract some useful values
```{r}

#How many replicates does this dataset contain?
nreps = length(replist)
print(sprintf("This dataset contains %s replicate simulations.", nreps))

#How many samples collected for each replicate? How many time samples maximum? (less when extinctions occur)
nsamples_list <- list()
for (i in 1:length(replist)){
  nsamples <-length(replist[[i]][[1]])
  nsamples_list[i] <- nsamples
}
maxsamples <- max(unlist(nsamples_list))
print(sprintf("At most, %s time samples were collected.", maxsamples))

#How many years of simulation does this represent?
nyears_list <- list()
for (i in 1:length(replist)){
  nyears <-replist[[i]][[1]][length(replist[[i]][[1]])]
  nyears_list[i] <- nyears
}

maxyears <- max(unlist(nyears_list))
print(sprintf("At most, this simulation was run for %s years.", maxyears))

#what years were sampled?
index_of_longest <- which.max(unlist(nyears_list))
sample_years <- replist[[index_of_longest]][[1]]
print("These years were sampled:")
print(unlist(sample_years))

```



## Plot some simple features of the total Rland population counts

```{r}
jpeg("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Rmetasim/replist_1_19_17_V6_seed_test_5000gen.jpg", width=600, height=400, units="px")

x1 <- replist[[1]][[1]] #list with values for generations sampled in replicate 1
y1 <- lapply(replist[[1]][[2]], function(x) {nrow(x)}) #list of matrices containing the data about individuals; the number of rows of each matrix gets counted to give number of individuals.

plot(x1,y1,xlim = c(0,maxyears), ylim=c(0,700000), main="Rland total inhabitants", sub = "5,000 years, 3 populations of 100,000 (98% seed), 4 age classes", xlab="Years", ylab="Total number of individuals")
lines(x1,y1)

for (i in 2:nreps) {
xs <- replist[[i]][[1]] #list with values for generations sampled in replicate
ys <- lapply(replist[[i]][[2]], function(x) {nrow(x)})
 # list with matrices; count length of each matrix
points(xs,ys)
lines(xs,ys)
}

#calculate the mean value for the 20 replicates
meanlist = list()
for (i in 1:maxsamples) { #loop over time
  values = list() #make an empty list to hold the values for all replicates at this timepoint
  for (j in 1:nreps) { #loop over replicates
    if(i %in% 1:length(replist[[j]][[1]]))
      {
      values[j] <- nrow(replist[[j]][[2]][[i]])
    } else {values[j] <- 0}
  }
  meanlist[i] <- mean(unlist(values))
}  


lines(sample_years, meanlist, col=rgb(1,0,0,alpha=0.6), lwd=3)
#points(sample_years, meanlist, col="red")

legend(0,680000, c("Replicate simulations", "Mean"), lty=c(1,1), pch=c(1,NA), lwd=c(1,3), col = c("black", rgb(1,0,0,alpha=0.6)))

dev.off()

```

This is first replicate and individual list contains all time points of rland individuals [1]: time scale
replist[[1]][[2]]
This is first replicate and list of generation times that coorespond to samples in individual list
replist[[1]][[1]]
```{r}
# at time 0
replist[[1]][[2]][1]

library(adegenet)

pops <- seq(0,4,2)

rland.df <- data.frame(replist[[2]][[2]][26])



alleles <- seq(7,29,2) #always start with 7, 29 = length of dataframe-1,

# add mapping of columns to what is in them 
# Combine alleles per locus to format for what genind wants
rland.2genind <- sapply(alleles, function(x){
  paste(rland.df[,x], rland.df[,x+1], sep = "-")
})

# find number of alleles over time - find when that is stable at what mutation rate? 

rland.df$pop <- "pop3"
rland.df$pop[rland.df$X1 < 8] <- "pop2"
rland.df$pop[rland.df$X1 < 4] <- "pop1"

#sample from 3, 7, and 11s
table(rland.df$X1)

frompops <- c(3,7,11)

rows <- lapply(frompops, function(x){
  n_pop <- length(rland.df$X1[rland.df$X1 == x])
  if(n_pop > 0){
    if(n_pop >= 30){
      sample(which(rland.df$X1 == x),30)
}   else {
      sample(which(rland.df$X1 == x),n_pop)
    }
  }
}
  )

rows <- do.call(c,rows)

samp.rland.2genind <- rland.2genind[rows,]
samp.rland.df <- rland.df[rows,]

library(strataG)

rland.genind <- df2genind(samp.rland.2genind, pop = samp.rland.df$pop, type = "codom", sep="-")
class(rland.genind)
rland.genind

#how many alleles per locus (average)
mean(rland.genind$loc.n.all)
#or the full list
rland.genind$loc.n.all
#


rland.gtypes <- genind2gtypes(rland.genind)

ot <- overallTest(rland.gtypes)
print(ot$result)
#print(ot$result[1,])

#print(ot$result[3,1])
#print(ot$result[3,2])


#data.class(ot$result)
#print(ot$strata.freq)


pws <- pairwiseTest(rland.gtypes)
#print("And try to sample")
#print(data.class(pws[1]$result))
pws[1]$result
#pws[1]$result[1,10]


```







## Structure of rland$individuals

From "Creating Landscapes" vignette, what the columns in the rland$individuals table mean
"The first six columns are always present and the last five shown here are a product of the choice of number of loci. The first column contains the individuals subpopulation and lifecycle stage (subpopulation = floor(x/rland\$intparam\$stages), lifecycle stage = x mod rland\$intparam\$stages). The second column is currently unused, always 0. The third column contains the generation in which the individual was born or created. The next three contain numerical ids for the individual, its mother and its father. After the first six columns the indivduals genetic code begins. The loci are shown in order with 2 columns for diploid loci and 1 column for haploid loci. The value of these columns represent the allele index of the allele the individual carries."






## Early drafts of other processing step


table of subpopulations
#```{r}


plot(replist[[1]][[1]], nrow(replist[[1]][[2]]))



numind <- sapply(rland.list[[2]], function(x){
  nrow(x)
})
subpops <- lapply(rland.list[[2]], function(x){
  table(x[,1])
})

subpops <- do.call(rbind, subpops)
head(subpops)

subpops <- data.frame(subpops)
subpops1<- melt(subpops, id.vars = rownames(subpops), value.name = "SubClass")

ggplot(subpops, aes())

#```



# Population Genetic statistics
#```{r}
library(adegenet)

pops <- seq(0,4,2)

rland.df <- as.data.frame(replist[[1]][[2]][1])

alleles <- seq(7,29,2)
rland.2genind <- sapply(alleles, function(x){
  paste(rland.df[,x], rland.df[,x+1], sep = "-")
})


rland.df$pop[rland.df$V1 == 0 | rland.df$V1 == 1 ] <- "pop1"
rland.df$pop[rland.df$V1 == 2 | rland.df$V1 == 3 ] <- "pop2"
rland.df$pop[rland.df$V1 == 4 | rland.df$V1 == 5 ] <- "pop3"

# following doesn't work yet!
rland.df$pop <- sapply(pops, function(x){
  pop <- if(rland.df$V1 == x | rland.df$V1 == x+1){
    paste("pop",x,sep="")
  }
  pop
})

str(rland.df)

rland.df2genind <- data.frame(rland.df[,c(3,31)], rland.2genind)

library(strataG)

head(rland.df2genind)

rland.genind <- df2genind(rland.df2genind[,-(1:2)], pop = rland.df2genind$pop, type = "codom", sep="-")
class(rland.genind)
rland.genind

rland.gtypes <- genind2gtypes(rland.genind)

overallTest(rland.gtypes)
pairwiseTest(rland.gtypes)



#```




