---
title: "Subsample Phacelia data to estimate sensitivity to sample size"
author: "Jessie Berta-Thompson"
date: "May 8, 2017"
output: html_document
---

Purpose of this version, immediate tasks:
 - subsample phacelia data to take samples of different sizes (2-30), then run strataG Fst. 
 - sample without replacement (real subsamples)


*************************************************************************************
Set up: clear history and make sure you're in a good working directory. 
Gather tools (reading and converting popgen data, running popgen calculations)
```{r}
rm(list=ls())
print("Where are we? Where will new files without specified full path be created?")
getwd()
library(ape)
library(phangorn)
library(ade4)
library(adegenet)
library(apex)
library(strataG)
```
*************************************************************************************
Step 1. Get the Phacelia microsatellite data

Read in the real data file (adegenet tools that take a genepop.gen file, make a genind object)
```{r}
real_data_by_loc.genind <-read.genepop("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Subsampling/Copy_of_real_data_PhfoAllGenePop_13Dec2016.gen")

#Note one error message: "Duplicate individual names detected. Coercing them to be unique." This is because .gen file has no individual names, just a list of genotypes, but genind requires individual names. They are given numbers, in order of original file.
```

Take a look at the genind object
```{r}
print("Real data genind (by sampling population)")
#print on whole object gives an overview
print(real_data_by_loc.genind)

#number of samples per location
print(table(real_data_by_loc.genind$pop))

```


Make 1000 randonly sampled datasets each for alternate sample sizes (2, 3, 4, 5, 6, 7, 8 ...28, 29, 30)

```{r}

sampling_options <- c(2:30) #various sample sizes to explore; 30 is boring - all samples will be the same.
bootreps <- 1000 # for each sample size, how many replicates datasets to create

#convert data into a dataframe for easy sampling (individual x locus, entries allele)
ind_locus_table.df <- genind2df(real_data_by_loc.genind, sep="-", pop = real_data_by_loc.genind$pop, usepop = FALSE)
#keep the population membership for each individual in a parallel list
poplist <- real_data_by_loc.genind$pop
#size of each sites' sampling effort as a vector
popcounts <- as.numeric(table(poplist))
#just the names of the sites (in same order as their counts above)
popnames <- names(table(poplist))
#total number of sites
npop <- length(popnames)


#loop over all the sizes of samples you want to make
sample_set <- lapply(sampling_options, function(samplen) {
  print(sprintf("Working on samples of %s", samplen))
  #loop over 1000 bootstrap replicates
  onethousanddatasets <- lapply(1:bootreps, function(bootrep) {
    #loop over 15 sites
    subsamples_by_site <- lapply(1:npop, function(p){
      #which site are we on?
      name <- popnames[p]
      #how many samples were originally taken there?
      n <- popcounts[p]
      if(n>=samplen) { # if there are enough real samples to subsample to desired level
        take <- samplen
      }
      else if (n<samplen){ # for the small number of samples with less than 30, just take max possible
        take <-n
      }
      #sample at random 'take' row indices, from the correct population.
      resampled_rows <- sample(which(poplist== name), take, replace = FALSE)
      #take rows from the data table and from population list based on sample
      subsetdata <- ind_locus_table.df[resampled_rows,]
      subsetpops <- poplist[resampled_rows]
      return(list(subsetdata, subsetpops)) #for each population, returns a list with 2 things. first is a data with some popgen data, second is a character vector of same length as number of rows in dataframe, containing population assignments. contains 1 sample dataset, sampling from all 15 populations.
    })

    #extract information output of list of lists constructed for this sample dataset
    resampleddatalist <- lapply(subsamples_by_site, function(sublist) {
      sublist[[1]]
    })
    resampledpoplist <- lapply(subsamples_by_site, function(sublist) {
      sublist[[2]]
    })
    

    resampled.df <- do.call("rbind", resampleddatalist) #make a list of dataframes into one dataframe
    resampledpops <- unlist(resampledpoplist) #make a list of lists into one vector

    #put back into a genind.
    reconvert.genind <- df2genind(resampled.df, sep="-", pop = resampledpops, ploidy=2, type="codom")
    print(sprintf("Done building subsampled dataset %s of %s for sample size n = %s", bootrep, bootreps, samplen))
    return(reconvert.genind) #output is one dataset, sampled at samplen level or max. 
})
#for each sample size, return many subsampled datasets
return(onethousanddatasets)
})

```

briefly check out product properties and slicing, then save

```{r}
print("Full Sample set")
print(data.class(sample_set))
print(length(sample_set))

print("One item in sample set: represents one sample size")
print(sprintf("Sample size n = %s",sampling_options[1]))
print(data.class(sample_set[[1]])) #need double brackets - weird list in list.
print(length(sample_set[[1]])) #this thing is a list of 1000 datasets

print("One item from the list of 1000 datasets for each sample size")
print(data.class(sample_set[[1]][[1]])) # a single genind object is accessed like this.
print(length(sample_set[[1]][[1]])) 
print(sample_set[[1]][[1]])
print("sampling scheme for n = 2, first replicate")
print(table(sample_set[[1]][[1]]$pop)) 

print("sampling scheme for n = 28, 736th replicate")
print(table(sample_set[[27]][[736]]$pop)) 

print("data for n = 4, first replicate")
print(sample_set[[3]][[1]]$tab[1:10, 1:7]) #10 individuals, looking at first 7 alleles 
print(sample_set[[3]][[500]]$pop[1:10])

print("data for n = 4, 500th replicate")
print(sample_set[[3]][[500]]$tab[1:10, 1:7])
print(sample_set[[3]][[500]]$pop[1:10])

#individuals' nicknames are still original, so you can see, ind. 1-16 are from HN. 


filepath <- "29x1000_geninds_subsampled_by_location_2-30n_5_8_17.RData" 
save(sample_set, file=filepath)
```

Then start popgen calculations.
```{r}
#For time's sake, only run Fst pairwise calculations.

t0 <- Sys.time() #start time for whole procedure
#loop over different sample sizes
testresults <- lapply(1:length(sampling_options), function(i){
  ts <- Sys.time() #start time for this n
  n <- sampling_options[i] #because samples go from 2-30, index is not n.
  #loop over replicate random sampling efforts
  all_reps_one_n <- lapply(1:bootreps, function(b) {
    one.genind <- sample_set[[i]][[b]]
    one.gtypes <- genind2gtypes(one.genind)
    pT <- pairwiseTest(one.gtypes, stats=c("Fst"), nrep=1000)
    tf <- Sys.time()
    print(sprintf("Done with rep %s of %s for n = %s, which took %s minutes. %s hours since start.", b, bootreps, n, format(as.numeric(tf-ts,units="mins"), digits=3), format(as.numeric(tf-t0,units="hours"), digits=3)))
  return(pT) #lapply over replicates returns a list of Fst results
  })
  return(all_reps_one_n) #lapply over sample sizes returns a list of lists of Fst results
})


filepath <- "29x1000_Fst_calcs_sample_size_2-30_by_location_5_8_17.RData" 
save(testresults, file=filepath)



```



















